### 2015-01-10 paper/paper6/paper6.tex
elsevier tex

%% JWS example using Elsevier elsart classes, adapted by Tim 
%% Finin, finin@cs.umbc.edu

%% comment out one of these to get preprint vs. journal format

%\documentclass{elsart}       %% one column for review, preprint
\documentclass{elsart3p}    %% two columns for publication

\usepackage{graphicx,amssymb}

\renewcommand\floatpagefraction{.2}
\makeatletter
\def\elsartstyle{%
    \def\normalsize{\@setfontsize\normalsize\@xiipt{14.5}}
    \def\small{\@setfontsize\small\@xipt{13.6}}
    \let\footnotesize=\small
    \def\large{\@setfontsize\large\@xivpt{18}}
    \def\Large{\@setfontsize\Large\@xviipt{22}}
    \skip\@mpfootins = 18\p@ \@plus 2\p@
    \normalsize
}
\@ifundefined{square}{}{\let\Box\square}
\makeatother

%% tell Latex where to look for files menioned in \includegraphics
\graphicspath{{jwsGraphics/}}

\begin{document}

\begin{frontmatter}

%% This TITLE has a note reference (titleNote) whose 
%% contants is specified about 10 lines later.

\title{RDF Validation of \\ Metadata on Highly-Complex Person-Level Data Sets}

%% Here are the three AUTHORS.  Note the separating commas
%% with no and.

\author[gesis]{Thomas Bosch\corauthref{C}},
\ead{thomas.bosch@gesis.org}
\author[gesis]{Benjamin Zapilko},
\ead{benjamin.zapilko@gesis.org}
\author[gesis]{Joachim Wackerow}
\ead{joachim.wackerow@gesis.org}
\author[mannheim]{Kai Eckert}
\ead{kai@informatik.uni-mannheim.de}

%% these specify any 'notes' for the title and authors,
%% including 'thanks' that acknowledge support, etc. and an
%% indication of who the corresponding author is if there's
%% more than one author.

%\thanks[titleNote]{Adapted from a document written by Simon Pepping}
%\thanks[fininGrants]{Partisl support provided by NSF award NSF-ITR-IIS-0326460} 
\corauth[C]{Corresponding author. Tel: + 49 (0) 621 / 1246-271}

\medskip

%% Here are the addresses refereenced for the authors.  If
%% all of the authors are from the same institution you need
%% not have the lables for the author and address commands.

\address[gesis]{Gesis - Leibniz Institute for the Social Sciences, 68159 Mannheim, Germany}
\address[mannheim]{University of Mannheim, 68159 Mannheim, Germany}

\begin{abstract}

\end{abstract}

%% you choose your own keywords

\begin{keyword}
 RDF Validation, Person-Level Metadata, DDI-RDF Discovery Vocabulary, Disco, Linked Data
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{introduction}

\bibliographystyle{elsart-num-sort}

\bibliography{../../literature/literature}

\end{document}