### 2015-03-24 paper/paper7/paper7-evaluation.tex
evaluation

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}

% allows for temporary adjustment of side margins
\usepackage{chngpage}

% just makes the table prettier (see \toprule, \bottomrule, etc. commands below)
\usepackage{booktabs}

\usepackage[utf8]{inputenc}
%\usepackage[font=small,skip=0pt]{caption}

% footnotes
\usepackage{scrextend}

% colors
\usepackage[usenames, dvipsnames]{color}

% underline
\usepackage{tikz}
\newcommand{\udensdot}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[densely dotted] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\uloosdot}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[loosely dotted] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\udash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\udensdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[densely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\uloosdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[loosely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

% URL handling
\usepackage{url}
\urlstyle{same}

%\usepackage{makeidx}  % allows for indexgeneration

%\usepackage{amsmath}
\usepackage{amsmath, amssymb}
\usepackage{mathabx}
\usepackage{caption} 
\captionsetup[table]{skip=10pt}

% monospace within text
\newcommand{\ms}[1]{\texttt{#1}}

% examples
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{ex}{Verbatim}{numbers=left,numbersep=2mm,frame=single,fontsize=\scriptsize}

\usepackage{xspace}
% Einfache und doppelte Anfuehrungszeichen
\newcommand{\qs}{``} 
\newcommand{\qe}{''\xspace} 
\newcommand{\sqs}{`} 
\newcommand{\sqe}{'\xspace}

% checkmark
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

% Xs
\usepackage{pifont}

% Tabellenabstände kleiner
\setlength{\intextsep}{10pt} % Vertical space above & below [h] floats
\setlength{\textfloatsep}{10pt} % Vertical space below (above) [t] ([b]) floats
% \setlength{\abovecaptionskip}{0pt}
% \setlength{\belowcaptionskip}{0pt}

\usepackage{tabularx}
\newcommand{\hr}{\hline\noalign{\smallskip}} % für die horizontalen linien in tabellen

% Todos
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\ke}[1]{\todo[size=\small, color=orange!40]{\textbf{Kai:} #1}}
\newcommand{\tb}[1]{\todo[size=\small, color=green!40]{\textbf{Thomas:} #1}}
\newcommand{\er}[1]{\todo[size=\small, color=red!40]{\textbf{Erman:} #1}}
\newcommand{\an}[1]{\todo[size=\small, color=blue!40]{\textbf{Andy:} #1}}

\newenvironment{table-1cols}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}
  \hline
  \textbf{Requirements} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{table-2cols}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|l}
  \hline
  \textbf{Requirements} & \textbf{Covering DSCLs} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{complexity}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|l}
  \hline
  \textbf{Complexity Class} & \textbf{Complexity} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{DL}{
  %\scriptsize
  %\sffamily
  \vspace{0cm}
  \begin{tabular}{r l}

}{
  \end{tabular}
  %\linebreak
}

\newenvironment{evaluation}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{Constraint Class} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{constraint-languages-complexity}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{Complexity Class} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{user-fiendliness}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c}
  \hline
  \textbf{criterion} & \textbf{DSP} & \textbf{OWL2} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\setcounter{secnumdepth}{5}

% tables
\usepackage{array,graphicx}
\usepackage{booktabs}
\usepackage{pifont}
\newcommand*\rot{\rotatebox{90}}
\newcommand*\OK{\ding{51}}
\usepackage{booktabs}
\newcommand*\ON[0]{$\surd$}

\usepackage{tablefootnote}

\usepackage{float}

\begin{document}
\renewcommand{\arraystretch}{1.3}
%
%
\title{An Evaluation of RDF Constraint Types to Validate Metadata on Person-Level and Aggregated Data}
\subtitle{}

\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Benjamin Zapilko\inst{1} \and Joachim Wackerow\inst{1} \and Kai Eckert\inst{2}}
%
\authorrunning{} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{firstname.lastname\}@gesis.org},\\ 
\and
University of Mannheim, Germany \\
\email{kai@informatik.uni-mannheim.de} 
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
...

\keywords{RDF Validation, RDF Constraints, DDI-RDF Discovery Vocabulary, Disco, RDF Data Cube Vocabulary, Linked Data, Semantic Web}
\end{abstract}

\section{Introduction}

\section{Evaluation}

\subsection{Legend}

\begin{table}
	\centering
		\begin{tabular}{c|l}
      \textbf{Symbol} & \textbf{Description} \\		
			\emph{X} & \textbf{Validation Successful} \\
			\hline
			\textbf{\textgreater} \emph{X} & \textbf{Poor Performance/Scaling} \\
			\hline
      \ding{55} & \textbf{Very Poor Performance/Scaling} \\
			\hline
			$\dagger$ & \textbf{SPARQL Endpoint Error}
		\end{tabular}
	\caption{Legend}
	\label{tab:legend}
\end{table}

\textbf{Validation Successful:} 
\emph{X} indicates the number of raised constraint violation triples.

\textbf{Poor Performance/Scaling:}
The performance of the implementation of the underlying SPARQL CONSTRUCT query 
is too poor to get all resulting constraint violation triples. 
Therefore, a limit of \emph{X} result constraint violation triples is set. 
It is likely that there are more than \emph{X} constraint violations.
Although, the result set contains not the whole set of raised constraint violation triples,
the constraint can be used as an indicator if there data not conforming to the constraint and
to resolve constraint violations step by step. 
As part of future work, the performance will be improved.

\textbf{Very Poor Performance/Scaling:} 
The performance of the implementation of the underlying SPARQL CONSTRUCT query 
is too poor to get any results, even though a limit of result constraint violation triples is set. 
As part of future work, the performance will be improved.

\section{Disco}

\begin{itemize}
	\item DwB Discovery Portal
	\begin{itemize}
	  \item Metadata records for each study are available on the study landing pages.
		\item \url{http://dwb-dev.nsd.uib.no/portal/#/study/http://portal.dwbproject.org/resource/Study/de.gesis/zacat/ZA5597/en}
    \item Search interface: \url{http://dwb-dev.nsd.uib.no/portal/#/search?q=smoking&p=1&s=score&d=desc&f=text}
	\end{itemize}
\end{itemize}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Disco Data Sets} \\		
      \hline
    Missy & Microdata Information System\tablefootnote{\url{http://www.gesis.org/missy/eu/missy-home}} \\
		DwB & DwB Discovery Portal\tablefootnote{\url{http://dwb-dev.nsd.uib.no/portal}} \\
		DDA-SND & DDA and SND DDI-RDF\tablefootnote{\url{http://ddi-rdf.borsna.se/}} \\ 
		\end{tabular}
	\caption{Disco Data Sets Abbreviations}
	\label{tab:disco-data-sets-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{9}{c}{\textbf{Counts}}
    \\  \cmidrule{2-10}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{disco:StudyGroup}}
           & \textbf{\rot{disco:Study}}
           & \textbf{\rot{disco:LogicalDataSet}}
           & \textbf{\rot{disco:Universe}}
					 & \textbf{\rot{disco:Variable}}
					 & \textbf{\rot{disco:Question}}
				   & \textbf{\rot{disco:SummaryStatistics}}
					 & \textbf{\rot{disco:CategoryStatistics}}
					 & \textbf{\rot{skos:Concept}}
    \\ \midrule
    Missy &  &  &  &  &  & \\
		DwB & 2,332,802 & 0 & 1,387 & 1,367 & 2,796 & 446,806 & 0 & 0 & 0 & 0 \\
		DDA-SND & 2,271,415 & 0 & 1,490 & 0 & 10,188 & 80,070 & 139,237 & 0 & 0 & 290,963 \\ 
    \bottomrule
    \end{tabular}
    \caption{Disco Data Sets Overview}
		\label{tab:disco-overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Data Sets} & \textbf{SPARQL Endpoint} \\		
      \hline
      Missy & \textcolor{red}{XXXXX} \\
			DwB & \url{http://dwb-dev.nsd.uib.no/sparql} \\
			DDA-SND & \url{http://ddi-rdf.borsna.se/endpoint/} \\
		\end{tabular}
	\caption{Disco SPARQL Endpoints}
	\label{tab:disco-sparql-endpoints}
\end{table}

\section{Data Cube}

\begin{itemize}
  \item overview over Data Cube data sets\footnote{\url{http://270a.info/}; \url{http://datahub.io/de/dataset?tags=format-qb}} 
  \item \url{http://ontologycentral.com/}
\end{itemize}

%\begin{table}
	%\centering
		%\begin{tabular}{l|c|c|c|c|c|c}
      %\textbf{Data Set} & \textbf{triples} & \textbf{data sets} & \textbf{data structure definitions} & \textbf{observations} & \textbf{slices} & \textbf{dimensions} \\		
      %\hline
      %European Central Bank (ECB)\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} &  &  &  &  &  &  \\
			%UNESCO Institute for Statistics (UIS)\tablefootnote{\url{http://www.uis.unesco.org/Pages/default.aspx}} &  &  &  &  &  &  \\
			%International Monetary Fund (IMF)\tablefootnote{\url{http://www.imf.org/external/index.htm}} &  &  &  &  &  &  \\
			%Bundesamt für Statistik - Swiss Federal Statistics (BFS)\tablefootnote{\url{http://www.bfs.admin.ch/}} &  &  &  &  &  &  \\
			%Food and Agriculture Organization of the United Nations (FAO)\tablefootnote{\url{http://www.fao.org/home/en/}} &  &  &  &  &  &  \\
			%World Bank (WB)\tablefootnote{\url{http://www.worldbank.org/}} &  &  &  &  &  &  \\
			%Federal Reserve Board (FRB)\tablefootnote{\url{http://www.federalreserve.gov/}} &  &  &  &  &  &  \\
			%Transparency International (TI)\tablefootnote{\url{http://www.transparency.org/}} &  &  &  &  &  &  \\
			%Organisation for Economic Co-operation and Development (OECD)\tablefootnote{\url{http://www.oecd.org/}} &  &  &  &  &  &  \\
			%Bank for International Settlements (BIS)\tablefootnote{\url{http://www.bis.org/}} &  &  &  &  &  &  \\
			%Australian Bureau of Statistics (ABS)\tablefootnote{\url{http://abs.gov.au/}} &  &  &  &  &  &  \\
		%\end{tabular}
	%\caption{Data Cube Data Sets}
	%\label{tab:data-cube-data-sets}
%\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Data Cube Data Sets} \\		
      \hline
    ECB & European Central Bank\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} \\
		UIS & UNESCO Institute for Statistics\tablefootnote{\url{http://www.uis.unesco.org/Pages/default.aspx}} \\
		IMF & International Monetary Fund\tablefootnote{\url{http://www.imf.org/external/index.htm}} \\
		BFS & Bundesamt für Statistik - Swiss Federal Statistics\tablefootnote{\url{http://www.bfs.admin.ch/}} \\
		FAO & Food and Agriculture Organization of the United Nations\tablefootnote{\url{http://www.fao.org/home/en/}} \\
		WB & World Bank\tablefootnote{\url{http://www.worldbank.org/}} \\
		FRB & Federal Reserve Board\tablefootnote{\url{http://www.federalreserve.gov/}} \\
		TI & Transparency International\tablefootnote{\url{http://www.transparency.org/}} \\
		OECD & Organisation for Economic Co-operation and Development\tablefootnote{\url{http://www.oecd.org/}} \\
		BIS & Bank for International Settlements\tablefootnote{\url{http://www.bis.org/}} \\
		ABS & Australian Bureau of Statistics\tablefootnote{\url{http://abs.gov.au/}} \\
		IEEE-VIS & IEEE VIS Source Data \\
		ACORN-SAT & Australian Climate Observations Reference Network - Surface Air Temperature Dataset \\
		HDP & HealthData.gov Platform (HDP) on the Semantic Web \\
		Eurostat & The Eurostat Linked Data  \\
		Asturias & Nomenclator Asturias \\
		ISTAT & ISTAT Immigration (LinkedOpenData.it) \\
		ICANE & Statistical Office of Cantabria (Instituto Cántabro de Estadística, ICANE) \\
		EE-2009 & European Election Results 2009 \\
		EU-B & Standard Eurobarometer \\
		ECB-S & European Central Bank Statistics (PublicData.eu) \\
		CPV-2008 & Common Procurement Vocabulary (CPV) 2008 \\
		CPV-2003 & Common Procurement Vocabulary (CPV) 2003 \\
		GHI-2011 & Global Hunger Index (GHI) 2011 \\
		\end{tabular}
	\caption{Data Cube Data Sets Abbreviations}
	\label{tab:data-cube-data-sets-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{5}{c}{\textbf{Counts}}
    \\  \cmidrule{2-6}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{qb:DataSet}}
           & \textbf{\rot{qb:DataStructureDefinition}}
           & \textbf{\rot{qb:Observation}}
           & \textbf{\rot{qb:Slice}}
    \\ \midrule
    ECB & 468,899,474 & 55 & 46 & $>$11,000,000 & 428,698 \\
		UIS & 10,400,534 & 5 & 5 & 1,437,651 & 0 &  \\
		IMF & 35,688,446 & 4 & 8 & 3,603,719 & 0 &  \\
		BFS & 1,533,743 & 0 & 0 & 8 & 0 &  \\
		FAO & 53,000,000 & 10 & 10 & $>$7,100,000 & 0 \\
		WB & 174,006,552 & 9,466 & 59 & $>$17,000,000 & 0 \\
		FRB & 185,266,900 & 49 & 98 & $>$9,500,000 & 0 \\
		TI & 52,233 & 6 & 6 & 3,928 & 0 \\
		OECD & 304,995,160 & 136 & 140 & $>$12,000,000 & 0 \\
		BIS & 54,197,482 & 6 & 12 & 3,606,466 & 47,914 \\
		ABS & 2,357,400,000 & 253 & 257 & $>$11,000,000 & 0  \\
		IEEE-VIS & 19,935,340 & 0 & 0 & 1,350 & 0 \\
		ACORN-SAT & 61,406,503 &  &  &  &  &  \\
		HDP & 1,934,046,908 &  &  &  &  &  \\
		Eurostat & 8,000,000,000 &  &  &  &  &  \\
		Asturias & 4,508,050 &  &  &  &  &  \\
		ISTAT & 3,024,396 &  &  &  &  &  \\
		ICANE & 52,412 &  &  &  &  &  \\
		EE-2009 & 3,165 &  &  &  &  &  \\
		EU-B & 1,193,494 &  &  &  &  &  \\
		ECB-S & 10,000,000 &  &  &  &  &  \\
		CPV-2008 & 803,311 &  &  &  &  &  \\
		CPV-2003 & 546,135 &  &  &  &  &  \\
		GHI-2011 & 19,369 (dump) &  &  &  &  &  \\
    \bottomrule
    \end{tabular}
    \caption{Data Cube Data Sets Overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Data Sets} & \textbf{SPARQL Endpoints} \\		
      \hline
      ECB & \url{http://ecb.270a.info/sparql} \\
			UIS & \url{http://uis.270a.info/sparql} \\
			IMF & \url{http://imf.270a.info/sparql} \\
			BFS & \url{http://bfs.270a.info/sparql} \\
			FAO & \url{http://fao.270a.info/sparql} \\
			WB & \url{http://worldbank.270a.info/sparql} \\
			FRB & \url{http://frb.270a.info/sparql} \\
			TI & \url{http://transparency.270a.info/sparql} \\
			OECD & \url{http://oecd.270a.info/sparql} \\
			BIS & \url{http://bis.270a.info/sparql} \\
			ABS & \url{http://abs.270a.info/sparql} \\
		\end{tabular}
	\caption{Data Cube Data Sets SPARQL Endpoints}
	\label{tab:data-cube-data-sets-sparql-endpoints}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Data Model Consistency}
           & \rot{ECB}
           & \rot{UIS}
           & \rot{IMF}
           & \rot{BFS}
           & \rot{FAO}
					 & \rot{WB}
					 & \rot{FRB}
					 & \rot{TI}
					 & \rot{OECD}
					 & \rot{BIS}
					 & \rot{ABS}
					 & \rot{IEEE-VIS}
    \\ \midrule
    DATA-MODEL-CONSISTENCY-01 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    DATA-MODEL-CONSISTENCY-02 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    DATA-MODEL-CONSISTENCY-03 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    DATA-MODEL-CONSISTENCY-04 & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ &  & & & $\checkmark$ \\
		DATA-MODEL-CONSISTENCY-05 &  &  &  &  &  & & & & & & & $\checkmark$ \\
		DATA-MODEL-CONSISTENCY-06 &  &  &  &  &  & \\
		DATA-MODEL-CONSISTENCY-07 &  &  &  &  &  & \\
		DATA-MODEL-CONSISTENCY-08 &  &  &  &  &  & \\
		DATA-MODEL-CONSISTENCY-09 &  &  &  &  &  & \\
		DATA-MODEL-CONSISTENCY-10 &  &  &  &  &  & \\
		DATA-MODEL-CONSISTENCY-11 &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Existential Quantifications}
           & \rot{ECB}
           & \rot{UIS}
           & \rot{IMF}
           & \rot{BFS}
           & \rot{FAO}
					 & \rot{WB}
					 & \rot{FRB}
					 & \rot{TI}
					 & \rot{OECD}
					 & \rot{BIS}
					 & \rot{ABS}
					 & \rot{IEEE-VIS}
    \\ \midrule
		EXISTENTIAL-QUANTIFICATIONS-01 & 9 & $\checkmark$ & 11 & 7 & 8 & 77 & 8 & 9 & 7 & 8 & 7 & $\checkmark$ \\
		EXISTENTIAL-QUANTIFICATIONS-02 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		EXISTENTIAL-QUANTIFICATIONS-03 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 59 & $\checkmark$ & 6 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		EXISTENTIAL-QUANTIFICATIONS-04 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Cardinality Restrictions}
           & \rot{ECB}
           & \rot{UIS}
           & \rot{IMF}
           & \rot{BFS}
           & \rot{FAO}
					 & \rot{WB}
					 & \rot{FRB}
					 & \rot{TI}
					 & \rot{OECD}
					 & \rot{BIS}
					 & \rot{ABS}
					 & \rot{IEEE-VIS}
    \\ \midrule
		MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01 &  &  &  &  &  & \\
		EXACT-UNQUALIFIED-CARDINALITY-RESTRICTIONS-01 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-01 & \ding{55} & 118 & 8 & 8 & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & 12 & \ding{55} & 1,350 \\
		EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-02 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Structure}
           & \rot{ECB}
           & \rot{UIS}
           & \rot{IMF}
           & \rot{BFS}
           & \rot{FAO}
					 & \rot{WB}
					 & \rot{FRB}
					 & \rot{TI}
					 & \rot{OECD}
					 & \rot{BIS}
					 & \rot{ABS}
					 & \rot{IEEE-VIS}
    \\ \midrule
		STRUCTURE-01 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		STRUCTURE-02 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Constraints}
           & \rot{ECB}
           & \rot{UIS}
           & \rot{IMF}
           & \rot{BFS}
           & \rot{FAO}
					 & \rot{WB}
					 & \rot{FRB}
					 & \rot{TI}
					 & \rot{OECD}
					 & \rot{BIS}
					 & \rot{ABS}
					 & \rot{IEEE-VIS}
    \\ \midrule
		PROPERTY-DOMAIN-01 &  &  &  &  &  & \\
		PROPERTY-RANGES-01 &  &  &  &  &  & \\
		DISJOINT-PROPERTIES-01 &  &  &  &  &  & \\
		DISJOINT-CLASSES-01 &  &  &  &  &  & \\
		EQUIVALENT-PROPERTIES-01 &  &  &  &  &  & \\
		UNIVERSAL-QUANTIFICATIONS-01 &  &  &  &  &  & \\
		MEMBERSHIP-IN-CONTROLLED-VOCABULARIES-01 &  &  &  &  &  & \\
		CONTEXT-SPECIFIC-VALID-CLASSES-01 &  &  &  &  &  & \\
		CONTEXT-SPECIFIC-VALID-PROPERTIES-01 &  &  &  &  &  & \\
		RECOMMENDED-PROPERTIES-01 &  &  &  &  &  & \\
		VALUE-IS-VALID-FOR-DATATYPE-01 &  &  &  &  &  & \\
		VOCABULARY-01 &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets}
    \end{center}
\end{table}

\section{SKOS}

\begin{itemize}
  \item overview over SKOS data sets\footnote{\url{http://datahub.io/de/dataset?tags=format-skos}}
	\item overview over thesauri\footnote{\url{http://datahub.io/de/dataset?tags=thesaurus}} (table \ref{tab:thesauri-overview})
\end{itemize}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Thesauri} \\		
      \hline
    TheSoz & Thesaurus for the Social Sciences\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} \\
	  STW & Thesaurus for Economics\tablefootnote{\url{http://zbw.eu/stw/versions/latest/about}} \\
	  AGROVOC & AGROVOC Multilingual agricultural thesaurus\tablefootnote{\url{http://202.45.139.84:10035/catalogs/fao/repositories/agrovoc}} \\
		UNESCO & UNESCO Thesaurus\tablefootnote{\url{http://skos.um.es/sparql/}} \\
		TGN & The Getty Thesaurus of Geographic Names\tablefootnote{\url{http://vocab.getty.edu/sparql}} \\
		EARTh & Environmental Applications Reference Thesaurus\tablefootnote{\url{http://linkeddata.ge.imati.cnr.it/resource/EARTh/}} \\
		ODT & Open Data Thesaurus\tablefootnote{\url{http://vocabulary.semantic-web.at/PoolParty/wiki/OpenData}} \\
		SLD & Spanish Linguistic Datasets\tablefootnote{\url{http://linguistic.linkeddata.es}} \\
		SSWT & Social Semantic Web Thesaurus\tablefootnote{\url{http://vocabulary.semantic-web.at/PoolParty/wiki/semweb}} \\
		GBA-GU & Thesaurus of the Geological Survey of Austria (GBA) - Geology Unit\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		GBA-GTS & Thesaurus of the Geological Survey of Austria (GBA) - Geologic Time Scale\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		GBA-L & Thesaurus of the Geological Survey of Austria (GBA) - Lithology\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		GBA-LU & Thesaurus of the Geological Survey of Austria (GBA) - Lithotectonic Unit\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		GEMET & GEneral Multilingual Environmental Thesaurus\tablefootnote{\url{http://www.eionet.europa.eu/gemet/}} \\
		EuroVoc & EuroVoc\tablefootnote{\url{http://open-data.europa.eu/de/data/dataset/eurovoc}} \\
		CECCT & Clean Energy and Climate Change Thesaurus\tablefootnote{\url{http://data.reegle.info/thesaurus/guide}} \\
		\end{tabular}
	\caption{Thesauri Abbreviations}
	\label{tab:thesauri-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{6}{c}{\textbf{Counts}}
    \\  \cmidrule{2-7}
    \\       \textbf{Thesauri}
           & \textbf{\rot{triples}}
           & \textbf{\rot{skos:ConceptScheme}}
           & \textbf{\rot{sko:Concept}}
           & \textbf{\rot{skos:broader}}
           & \textbf{\rot{skos:narrower}}
					 & \textbf{\rot{skos:hasTopConcept}}
					 & \textbf{\rot{skos:inScheme}}
    \\ \midrule
    TheSoz & 439,153 & 1 & 8,426 & 13,705 & 13,706 & 0 & 48,529 \\
	  STW & 221,668 & 1 & 13,468 & 13,732 & 13732 & 7 & 13,180 \\
	  AGROVOC & 6,080,477 & 1 & 32,310 & 33,507 & 33,507 & 25 & 32,310 \\
		UNESCO & 288,346 & 9 & 26,714 & 20,028 & 20,028 & 607 & 32,009 \\
		TGN & 16,112,321 & 8 & 2,898,775 & 0 & 0 & 0 & 1,453,767\\
		EARTh & 9,287,364 &  &  &  & \\
		ODT & 3,290 &  &  &  & \\
		SLD & 7,629,211 &  &  &  & \\
		SSWT & 64,698 &  &  &  & \\
		GBA-GU & 25,718 &  &  &  & \\
		GBA-GTS & 7,875 &  &  &  & \\
		GBA-L & 9,317 &  &  &  & \\
		GBA-LU & 9,504 &  &  &  & \\
		GEMET & 372,889,229 &  &  &  & \\
		EuroVoc & 64,477,774 &  &  &  & \\
		CECCT & 191,336 &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Overview}
		\label{tab:thesauri-overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Thesauri} & \textbf{SPARQL Endpoints} \\		
      \hline
      TheSoz & \url{http://lod.gesis.org/thesoz/sparql} \\
			STW & \url{http://zbw.eu/beta/sparql/stw/query} \\
			AGROVOC & \url{http://202.45.139.84:10035/catalogs/fao/repositories/agrovoc} \\
			UNESCO & \url{http://skos.um.es/sparql/} \\
			TGN & \url{http://vocab.getty.edu/} \\
			EARTh & \url{http://linkeddata.ge.imati.cnr.it:8890/sparql} \\
			ODT & \url{http://vocabulary.semantic-web.at/PoolParty/sparql/OpenData} \\
			SLD & \url{http://linguistic.linkeddata.es/sparql} \\
			SSWT & \url{http://vocabulary.semantic-web.at/PoolParty/sparql/semweb} \\
			GBA-GU & \url{http://resource.geolba.ac.at/PoolParty/sparql/GeologicUnit} \\
			GBA-GTS & \url{http://resource.geolba.ac.at/PoolParty/sparql/GeologicTimeScale} \\
			GBA-L & \url{http://resource.geolba.ac.at/PoolParty/sparql/lithology} \\
			GBA-LU & \url{http://resource.geolba.ac.at/PoolParty/sparql/tectonicunit} \\
			GEMET & \url{http://semantic.eea.europa.eu/sparql} \\
			EuroVoc & \url{http://open-data.europa.eu/de/linked-data} \\
			CECCT & \url{http://poolparty.reegle.info/PoolParty/sparql/glossary}
		\end{tabular}
	\caption{Thesauri SPARQL Endpoints}
	\label{tab:thesauri-sparql-endpoints}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Data Model Consistency}
           & \rot{TheSoz}
           & \rot{STW}
           & \rot{AGROVOC}
					 & \rot{TGN}
           & \rot{UNESCO}
					 & \rot{ODT}
					 & \rot{SSWT}
					 & \rot{GBA-GU}
					 & \rot{GBA-GTS}
					 & \rot{GBA-L}
					 & \rot{GBA-LU}
					 & \rot{CECCT}
    \\ \midrule
    DATA-MODEL-CONSISTENCY-01 &  &  &  &  &  & \\
    DATA-MODEL-CONSISTENCY-02 &  &  &  &  &  & \\
    DATA-MODEL-CONSISTENCY-03 &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation}
		\label{tab:thesauri-evaluation-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Labeling and Documentation}
           & \rot{TheSoz}
           & \rot{STW}
           & \rot{AGROVOC}
					 & \rot{TGN}
           & \rot{UNESCO}
					 & \rot{ODT}
					 & \rot{SSWT}
					 & \rot{GBA-GU}
					 & \rot{GBA-GTS}
					 & \rot{GBA-L}
					 & \rot{GBA-LU}
					 & \rot{CECCT}
    \\ \midrule
		LABELING-AND-DOCUMENTATION-01 & 8,426 & 11,508 & 19,829 & $\checkmark$ & & 36 & 1,475 & 6 & 2 & $\checkmark$ & 107 & 486 \\
		LABELING-AND-DOCUMENTATION-02 & $>$1 & \ding{55} & $>$100 & $\checkmark$ & & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		LABELING-AND-DOCUMENTATION-03 & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ \\
		LABELING-AND-DOCUMENTATION-04 &  &  &  &  &  & & &  \\
		LABELING-AND-DOCUMENTATION-05 & $\checkmark$ & $\checkmark$ & \ding{55} & $\checkmark$ & & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & 3 \\
		LABELING-AND-DOCUMENTATION-06 & $>$1 & \ding{55} & \ding{55} & $\checkmark$ & & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation}
		\label{tab:thesauri-evaluation-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Structure}
           & \rot{TheSoz}
           & \rot{STW}
           & \rot{AGROVOC}
					 & \rot{TGN}
           & \rot{UNESCO}
					 & \rot{ODT}
					 & \rot{SSWT}
					 & \rot{GBA-GU}
					 & \rot{GBA-GTS}
					 & \rot{GBA-L}
					 & \rot{GBA-LU}
					 & \rot{CECCT}
    \\ \midrule
		STRUCTURE-01 & 1 & 1,074 & $\checkmark$ & $\checkmark$ &  & 5 & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		STRUCTURE-02 &  &  &  &  &  & \\
		STRUCTURE-03 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		STRUCTURE-04 & 2,906 & $>$1,000 & 726 & $\checkmark$ &  & 12 & 124 & 84 & 256 & $>$54 & 22 & 2,422 \\
		STRUCTURE-05 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ &  & 90 & 5,150 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 9,864 \\
		STRUCTURE-06 & 2 & $\checkmark$ & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 113 \\
		STRUCTURE-07 & 8 & 1,074 & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		STRUCTURE-08 &  &  &  &  &  & \\
		STRUCTURE-09 & 6,741 & $>$3,297 & 135 & $\checkmark$ &  & 2 & 16 & 25 & $\checkmark$ & $\checkmark$ & $\checkmark$ & 81 \\
		STRUCTURE-10 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation}
		\label{tab:thesauri-evaluation-3}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Language Tag Cardinality}
           & \rot{TheSoz}
           & \rot{STW}
           & \rot{AGROVOC}
					 & \rot{TGN}
           & \rot{UNESCO}
					 & \rot{ODT}
					 & \rot{SSWT}
					 & \rot{GBA-GU}
					 & \rot{GBA-GTS}
					 & \rot{GBA-L}
					 & \rot{GBA-LU}
					 & \rot{CECCT}
    \\ \midrule
		LANGUAGE-TAG-CARDINALITY-01 & 9,435 & 13,468 & 98,894 & $\checkmark$ &  & 541 & 10,147 & 5,117 & 2,061 & 1,742 & 2,272 & 15,550 \\
		LANGUAGE-TAG-CARDINALITY-02 & 8,222 & 36,936 & \ding{55} & $\checkmark$ &  & 265 & 3,627 & 2,212 & 635 & 631 & 1,253 & 9,607 \\
		LANGUAGE-TAG-CARDINALITY-03 & 8,222 & $\checkmark$ & 135 & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		LANGUAGE-TAG-CARDINALITY-04 & $\checkmark$ & 476 & \ding{55} & 50 &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation}
		\label{tab:thesauri-evaluation-4}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Constraints}
           & \rot{TheSoz}
           & \rot{STW}
           & \rot{AGROVOC}
					 & \rot{TGN}
           & \rot{UNESCO}
					 & \rot{ODT}
					 & \rot{SSWT}
					 & \rot{GBA-GU}
					 & \rot{GBA-GTS}
					 & \rot{GBA-L}
					 & \rot{GBA-LU}
					 & \rot{CECCT}
    \\ \midrule
		PROPERTY-DOMAIN-01 &  &  &  &  &  & \\
		PROPERTY-RANGES-01 &  &  &  &  &  & \\
		DISJOINT-PROPERTIES-01 &  &  &  &  &  & \\
		DISJOINT-PROPERTIES-02 &  &  &  &  &  & \\
		DISJOINT-CLASSES-01 &  &  &  &  &  & \\
		EQUIVALENT-PROPERTIES-01 &  &  &  &  &  & \\
		UNIVERSAL-QUANTIFICATIONS-01 &  &  &  &  &  & \\
		CONTEXT-SPECIFIC-VALID-CLASSES-01 &  &  &  &  &  & \\
		CONTEXT-SPECIFIC-VALID-PROPERTIES-01 &  &  &  &  &  & \\
		RECOMMENDED-PROPERTIES-01 &  &  &  &  &  & \\
		VOCABULARY-01 &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation}
		\label{tab:thesauri-evaluation-5}
    \end{center}
\end{table}

\section{XKOS}

\begin{itemize}
	\item Nomenclature d'activités française (NAF): French classification expressed in XKOS. the French refinement of the NACE, because it has explanatory notes.
	\item Nomenclature des Professions et Catégories Socioprofessionnelles (PCS): French classification expressed in XKOS.
	\item Nomenclature des catégories juridiques (CJ): French classification expressed in XKOS.
	\item ISIC: has explanatory notes too. 
	\item ISCO
\end{itemize}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{6}{c}{\textbf{Counts}}
    \\  \cmidrule{2-7}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
					 & \textbf{\rot{}}
					 & \textbf{\rot{}}
    \\ \midrule
    Nomenclature d'activités française (NAF)\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} &  &  &  &  &  & \\
		Nomenclature des Professions et Catégories Socioprofessionnelles (PCS)\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} &  &  &  &  &  & \\
		Nomenclature des catégories juridiques (CJ)\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} &  &  &  &  &  & \\
		ISIC &  &  &  &  &  & \\
		ISCO &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Statistical Classifications Overview}
		\label{tab:statistical-classifications-overview}
    \end{center}
\end{table}

\bibliography{../../literature/literature}{}
\bibliographystyle{plain}
\setcounter{tocdepth}{1}
%\listoftodos
\end{document}