### 2015-04-08 paper/paper7/paper7-evaluation.tex
evaluation

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}

% allows for temporary adjustment of side margins
\usepackage{chngpage}

% just makes the table prettier (see \toprule, \bottomrule, etc. commands below)
\usepackage{booktabs}

\usepackage[utf8]{inputenc}
%\usepackage[font=small,skip=0pt]{caption}

% footnotes
\usepackage{scrextend}

% colors
\usepackage[usenames, dvipsnames]{color}

% underline
\usepackage{tikz}
\newcommand{\udensdot}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[densely dotted] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\uloosdot}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[loosely dotted] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\udash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\udensdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[densely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

\newcommand{\uloosdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {#1};
        \draw[loosely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

% URL handling
\usepackage{url}
\urlstyle{same}

%\usepackage{makeidx}  % allows for indexgeneration

%\usepackage{amsmath}
\usepackage{amsmath, amssymb}
\usepackage{mathabx}
\usepackage{caption} 
\captionsetup[table]{skip=10pt}

% monospace within text
\newcommand{\ms}[1]{\texttt{#1}}

% examples
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{ex}{Verbatim}{numbers=left,numbersep=2mm,frame=single,fontsize=\scriptsize}

\usepackage{xspace}
% Einfache und doppelte Anfuehrungszeichen
\newcommand{\qs}{``} 
\newcommand{\qe}{''\xspace} 
\newcommand{\sqs}{`} 
\newcommand{\sqe}{'\xspace}

% checkmark
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

% Xs
\usepackage{pifont}

% Tabellenabstände kleiner
\setlength{\intextsep}{10pt} % Vertical space above & below [h] floats
\setlength{\textfloatsep}{10pt} % Vertical space below (above) [t] ([b]) floats
% \setlength{\abovecaptionskip}{0pt}
% \setlength{\belowcaptionskip}{0pt}

\usepackage{tabularx}
\newcommand{\hr}{\hline\noalign{\smallskip}} % für die horizontalen linien in tabellen

% Todos
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\ke}[1]{\todo[size=\small, color=orange!40]{\textbf{Kai:} #1}}
\newcommand{\tb}[1]{\todo[size=\small, color=green!40]{\textbf{Thomas:} #1}}
\newcommand{\er}[1]{\todo[size=\small, color=red!40]{\textbf{Erman:} #1}}
\newcommand{\an}[1]{\todo[size=\small, color=blue!40]{\textbf{Andy:} #1}}

\newenvironment{table-1cols}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}
  \hline
  \textbf{Requirements} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{table-2cols}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|l}
  \hline
  \textbf{Requirements} & \textbf{Covering DSCLs} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{complexity}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|l}
  \hline
  \textbf{Complexity Class} & \textbf{Complexity} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{DL}{
  %\scriptsize
  %\sffamily
  \vspace{0cm}
  \begin{tabular}{r l}

}{
  \end{tabular}
  %\linebreak
}

\newenvironment{evaluation}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{Constraint Class} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{constraint-languages-complexity}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{Complexity Class} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\newenvironment{user-fiendliness}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c}
  \hline
  \textbf{criterion} & \textbf{DSP} & \textbf{OWL2} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
}

\setcounter{secnumdepth}{5}

% tables
\usepackage{array,graphicx}
\usepackage{booktabs}
\usepackage{pifont}
\newcommand*\rot{\rotatebox{90}}
\newcommand*\OK{\ding{51}}
\usepackage{booktabs}
\newcommand*\ON[0]{$\surd$}

\usepackage{tablefootnote}

\usepackage{float}

\begin{document}
\renewcommand{\arraystretch}{1.3}
%
%
\title{An Evaluation of RDF Constraint Types to Validate Metadata on Person-Level and Aggregated Data}
\subtitle{}

\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Benjamin Zapilko\inst{1} \and Joachim Wackerow\inst{1} \and Kai Eckert\inst{2}}
%
\authorrunning{} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{\{firstname.lastname\}@gesis.org},\\ 
\and
University of Mannheim, Germany \\
\email{kai@informatik.uni-mannheim.de} 
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
...

\keywords{RDF Validation, RDF Constraints, DDI-RDF Discovery Vocabulary, Disco, RDF Data Cube Vocabulary, Linked Data, Semantic Web}
\end{abstract}

\section{Evaluation}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccc@{}}
           & \multicolumn{6}{c}{\textbf{Vocabularies}}
    \\  \cmidrule{2-7}
    \\       \textbf{Criteria (Counts)}
           & \textbf{\rot{Person-Level Metadata (Disco)}}
           & \textbf{\rot{Aggregated Metadata (Data Cube)}}
					 & \textbf{\rot{Thesauri (SKOS)}}
					 & \textbf{\rot{Rectangular Data (PHDD)}}
					 & \textbf{\rot{Statistical Classifications (XKOS)}}
					 & \textbf{\rot{Total}}
    \\ \midrule
		\emph{Validated Triples} & 9,673,055 & 3,775,983,610 & 477,737,281 \\
		\emph{Validated Data Sets} & 1,526 & 9,990 & 4,178 \\
		\hline
		\emph{Constraint Violations} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{0}$)} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{1}$)} & 473,574 $\equiv$ \textcolor{red}{XXXXX}\% \\
    \emph{Constraint Violations ($\mathcal{SL}_{2}$)} & \textcolor{red}{XXXXX} \\
				\hline
		\emph{Constraint Types} & 52 (15\textbar 37) \\
		\emph{Constraint Types} ($\mathcal{C}_{C}$) & 30 $\equiv$ 57.7 \% \\
		\emph{Constraint Types} ($\mathcal{C}_{T}$) & 22 $\equiv$ 42.3 \% \\
		\hline
		\emph{Constraints} & 142 (77\textbar 65) \\
		\emph{Constraints} ($\mathcal{C}_{C}$) & 72 $\equiv$ 50.7 \% \\
		\emph{Constraints} ($\mathcal{C}_{T}$) & 70 $\equiv$ 49.3 \% \\
		\emph{Constraints ($\mathcal{SL}_{0}$)} & 75 $\equiv$ 52.8\% (44\textbar 31)\\
		\emph{Constraints ($\mathcal{SL}_{1}$)} & 9 $\equiv$ 6.3\% (8\textbar 1)\\
		\emph{Constraints ($\mathcal{SL}_{2}$)} & 58 $\equiv$ 40.8\% (25\textbar 33)\\
    \bottomrule
    \end{tabular}
    \caption{Evaluation}
		\label{tab:evaluation}
    \end{center}
\end{table}

\subsection{Legend}

\begin{table}
	\centering
		\begin{tabular}{c|l}
      \textbf{Symbol} & \textbf{Description} \\		
			\textbf{\emph{X}} & Validation Successful \\
			\hline
			\textbf{\textgreater \emph{X}} & Poor Performance/Scaling \\
			\hline
      \ding{55} & Very Poor Performance/Scaling \\
			\hline
			\textbf{\emph{(!)}} & Not Yet Implemented Constraint \\
			\hline
			\textbf{\emph{(X)}} & The validation of \emph{X} data sets could not be finished \\
			           & due to SPARQL endpoints' technical restrictions (e.g. the defined timeout). \\
			\hline
			\textbf{\emph{\textsuperscript{*}}} & severity level $\mathcal{SL}_{0}$ (informational) \\
			\hline
			\textbf{\emph{\textsuperscript{**}}} & severity level $\mathcal{SL}_{1}$ (warning) \\
			\hline
			\textbf{\emph{\textsuperscript{***}}} & severity level $\mathcal{SL}_{2}$ (error) \\
		\end{tabular}
	\caption{Legend}
	\label{tab:legend}
\end{table}

\textbf{Validation Successful:} 
\emph{X} indicates the number of raised constraint violation triples.

\textbf{Poor Performance/Scaling:}
The performance of the implementation of the underlying SPARQL CONSTRUCT query 
is too poor to get all resulting constraint violation triples. 
Therefore, a limit of \emph{X} result constraint violation triples is set. 
It is likely that there are more than \emph{X} constraint violations.
Although, the result set contains not the whole set of raised constraint violation triples,
the constraint can be used as an indicator if there data not conforming to the constraint and
to resolve constraint violations step by step. 
As part of future work, the performance will be improved.

\textbf{Very Poor Performance/Scaling:} 
The performance of the implementation of the underlying SPARQL CONSTRUCT query 
is too poor to get any results, even though a limit of result constraint violation triples is set. 
As part of future work, the performance will be improved.

\section{Person-Level Metadata (Disco)}

\subsection{Evaluation Results}

Legend: total (implemented\textbar not yet implemented)

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Evaluation Criteria} & \textbf{Counts} \\		
      \hline
		\emph{Validated Triples} & 9,673,055 \\
		\emph{Validated Data Sets} & 1,526 \\
		\hline
		\emph{Constraint Violations} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{0}$)} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{1}$)} & 473,574 $\equiv$ \textcolor{red}{XXXXX}\% \\
    \emph{Constraint Violations ($\mathcal{SL}_{2}$)} & \textcolor{red}{XXXXX} \\
		\hline
		\emph{Constraint (Most Constraint Violations)} & \emph{DISCO-C-LABELING-AND-DOCUMENTATION-06} (547,916) \\
		                                               & \emph{DISCO-C-COMPARISON-VARIABLES-02} (547,916) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{0}$))} & \emph{DISCO-C-LABELING-AND-DOCUMENTATION-06} (547,916) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{1}$))} & \emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-46} (468,807) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{2}$))} & \emph{DISCO-C-COMPARISON-VARIABLES-02} (547,916) \\
		\hline
		\emph{Constraint Types} & 52 (15\textbar 37) \\
		\emph{Constraint Types} ($\mathcal{C}_{C}$) & 30 $\equiv$ 57.7 \% \\
		\emph{Constraint Types} ($\mathcal{C}_{T}$) & 22 $\equiv$ 42.3 \% \\
		\hline
		\emph{Constraint Types (Most Constraints)} & 1. Existential Quantifications: 46 $\equiv$ 32.4\% (46\textbar 0) \\
		                                            & 2. Data Model Consistency: 7 (1\textbar 6) \\
																								& 3. Aggregation: 7 (0\textbar 7) \\
		\emph{Constraint Type (Most Constraints ($\mathcal{SL}_{2}$))} & Existential Quantifications: 9 (9\textbar 0) \\
		\hline
		\emph{Constraints} & 142 (77\textbar 65) \\
		\emph{Constraints} ($\mathcal{C}_{C}$) & 72 $\equiv$ 50.7 \% \\
		\emph{Constraints} ($\mathcal{C}_{T}$) & 70 $\equiv$ 49.3 \% \\
		\emph{Constraints ($\mathcal{SL}_{0}$)} & 75 $\equiv$ 52.8\% (44\textbar 31)\\
		\emph{Constraints ($\mathcal{SL}_{1}$)} & 9 $\equiv$ 6.3\% (8\textbar 1)\\
		\emph{Constraints ($\mathcal{SL}_{2}$)} & 58 $\equiv$ 40.8\% (25\textbar 33)\\
		\end{tabular}
    \caption{Evaluation of Disco Data Sets - Evaluation Results}
		\label{tab:disco-evaluation-results}
\end{table}

%\begin{table}[H]
    %\begin{center}
    %\begin{tabular}{@{}lc@{}}
           %& \multicolumn{1}{c}{\textbf{Constraint Types}}
    %\\  \cmidrule{2-2}
    %\\       \textbf{Criteria (Counts)}
           %& \textbf{\rot{Existential Quantifications}}
    %\\ \midrule
		%\emph{Constraints / Constraint Type} & 46 \\
    %\bottomrule
    %\end{tabular}
    %\caption{Evaluation of Disco Data Sets - Evaluation Results (2)}
		%\label{tab:disco-evaluation-results-2}
    %\end{center}
%\end{table}

\subsection{Data Sets Overview}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Disco Data Sets} \\		
      \hline
    \emph{Missy} & \emph{Microdata Information System}\tablefootnote{\url{http://www.gesis.org/missy/eu/missy-home}} \\
		\emph{DwB} & \emph{DwB Discovery Portal}\tablefootnote{\url{http://dwb-dev.nsd.uib.no/portal}} \\
		\emph{DDA-SND} & \emph{DDI-RDF}\tablefootnote{\url{http://ddi-rdf.borsna.se/}} \\
		               & provided by the \emph{Danish Data Archive (DDA)}\tablefootnote{\url{http://samfund.dda.dk/dda/default-en.asp}} and Swedish National Data Service (SND)\tablefootnote{\url{http://snd.gu.se/en}} \\ 
		\end{tabular}
	\caption{Disco Data Sets Abbreviations}
	\label{tab:disco-data-sets-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{9}{c}{\textbf{Counts}}
    \\  \cmidrule{2-10}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{disco:StudyGroup}}
           & \textbf{\rot{disco:Study}}
           & \textbf{\rot{disco:LogicalDataSet}}
           & \textbf{\rot{disco:Universe}}
					 & \textbf{\rot{disco:Variable}}
					 & \textbf{\rot{disco:Question}}
				   & \textbf{\rot{disco:SummaryStatistics}}
					 & \textbf{\rot{disco:CategoryStatistics}}
					 & \textbf{\rot{skos:Concept}}
    \\ \midrule
    \emph{Missy} & 5,068,838 & 6 & 45 & 159 & 1,125 & 21,040 & 0 & 0 & 0 & 147,193 \\
		\emph{DwB} & 2,332,802 & 0 & 1,387 & 1,367 & 2,796 & 446,806 & 0 & 0 & 0 & 0 \\
		\emph{DDA-SND} & 2,271,415 & 0 & 1,490 & 0 & 10,188 & 80,070 & 139,237 & 0 & 0 & 290,963 \\ 
		\hline
		\textbf{\emph{Total}} & 9,673,055 & & & 1,526 \\
    \bottomrule
    \end{tabular}
    \caption{Disco Data Sets Overview}
		\label{tab:disco-overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Data Sets} & \textbf{SPARQL Endpoint} \\		
      \hline
      \emph{Missy} & \url{http://svko-missy:8181/openrdf-workbench/repositories/native-java-store/summary} \\
			\emph{DwB} & \url{http://dwb-dev.nsd.uib.no/sparql} \\
			\emph{DDA-SND} & \url{http://ddi-rdf.borsna.se/endpoint/} \\
		\end{tabular}
	\caption{Disco SPARQL Endpoints}
	\label{tab:disco-sparql-endpoints}
\end{table}

\subsection{Detailed Evaluation}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Existential Quantifications (1)}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-02}\textsuperscript{***} & 7 & 17 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-03}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-04}\textsuperscript{*} & 11,021 & 445,381 & 62,260 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-05}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & 139,237 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-06}\textsuperscript{*} & 12 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-07}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-08}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-09}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-10}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Existential Quantifications (1)}
		\label{tab:evaluation-disco-existential-quantifications-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Existential Quantifications (2)}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-11}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-12}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-13}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-14}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-15}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-16}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-17}\textsuperscript{*} & 159 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-18}\textsuperscript{*} & 159 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-19}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-20}\textsuperscript{*} & $\checkmark$ & 1,367 & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Existential Quantifications (2)}
		\label{tab:evaluation-disco-existential-quantifications-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Existential Quantifications (3)}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-21}\textsuperscript{*} & $\checkmark$ & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-22}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-23}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-24}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-25}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-26}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-27}\textsuperscript{***} & $\checkmark$ & 130 & 1,490 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-28}\textsuperscript{**} & 159 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-29}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-30}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Existential Quantifications (3)}
		\label{tab:evaluation-disco-existential-quantifications-3}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Existential Quantifications (4)}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-31}\textsuperscript{**} & 159 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-32}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-33}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-34}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-35}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-36}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-37}\textsuperscript{*} & 18,625 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-38}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & 750 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-39}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-40}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & 139,237 \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Existential Quantifications (4)}
		\label{tab:evaluation-disco-existential-quantifications-4}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Existential Quantifications (5)}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-41}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-42}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-43}\textsuperscript{*} & 15,733 & 446,806 & 80,070 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-44}\textsuperscript{*} & 159 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-45}\textsuperscript{*} & 6,784 & 446,806 & 19,221 \\
		\emph{DISCO-C-EXISTENTIAL-QUANTIFICATIONS-46}\textsuperscript{**} & 11,550 & 446,806 & 10,451 \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Existential Quantifications (5)}
		\label{tab:evaluation-disco-existential-quantifications-5}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Conditional Properties}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{ DISCO-C-CONDITIONAL-PROPERTIES-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & 80,070 \\
		\emph{ DISCO-C-CONDITIONAL-PROPERTIES-02}\textsuperscript{**} & 12 & $\checkmark$ & $\checkmark$ \\
		\emph{ DISCO-C-CONDITIONAL-PROPERTIES-03}\textsuperscript{**} & 90 & $\checkmark$ & 2,980 \\
		\emph{ DISCO-C-CONDITIONAL-PROPERTIES-04}\textsuperscript{***} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{ DISCO-C-CONDITIONAL-PROPERTIES-05}\textsuperscript{***} & 45 & 1,387 & 1,490 \\
		\emph{ DISCO-C-CONDITIONAL-PROPERTIES-06}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Conditional Properties}
		\label{tab:evaluation-disco-conditional-properties}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Provenance}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-PROVENANCE-01}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-PROVENANCE-02}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-PROVENANCE-03}\textsuperscript{*} & 159 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-PROVENANCE-04}\textsuperscript{*} & $\checkmark$ & 1,367 & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Provenance}
		\label{tab:evaluation-disco-provenance}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Labeling and Documentation}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-LABELING-AND-DOCUMENTATION-01}\textsuperscript{*} & 6 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-LABELING-AND-DOCUMENTATION-02}\textsuperscript{*} & 45 & 1,387 & 1,490 \\
		\emph{DISCO-C-LABELING-AND-DOCUMENTATION-03}\textsuperscript{*} & 159 & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-LABELING-AND-DOCUMENTATION-04}\textsuperscript{*} & $\checkmark$ & 1,367 & $\checkmark$ \\
		\emph{DISCO-C-LABELING-AND-DOCUMENTATION-05}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-LABELING-AND-DOCUMENTATION-06}\textsuperscript{*} & 21,040 & 446,806 & 80,070 \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Labeling and Documentation}
		\label{tab:evaluation-disco-labeling-and-documentation}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Data Model Consistency}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-DATA-MODEL-CONSISTENCY-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-02 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-03 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-04 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-05}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-06 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DATA-MODEL-CONSISTENCY-07 (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Data Model Consistency}
		\label{tab:evaluation-disco-data-model-consistency}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Comparison}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-COMPARISON-VARIABLES-01 (!)}\textsuperscript{**} \\
		\emph{DISCO-C-COMPARISON-VARIABLES-02}\textsuperscript{***} & 21,040 & 446,806 & 80,070 \\
		\emph{DISCO-C-COMPARISON-VARIABLES-03 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-COMPARISON-VARIABLES-04}\textsuperscript{*} & 18,625 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-COMPARISON-VARIABLES-05}\textsuperscript{***} & 159 & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Comparison}
		\label{tab:evaluation-disco-comparison}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Mathematical Operations}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-MATHEMATICAL-OPERATIONS-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-MATHEMATICAL-OPERATIONS-02 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-MATHEMATICAL-OPERATIONS-03 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-MATHEMATICAL-OPERATIONS-04 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-MATHEMATICAL-OPERATIONS-05 (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Mathematical Operations}
		\label{tab:evaluation-disco-mathematical-operations}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Language Tags}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-LANGUAGE-TAG-MATCHING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-LANGUAGE-TAG-CARDINALITY-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-LANGUAGE-TAG-CARDINALITY-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-LANGUAGE-TAG-CARDINALITY-03 (!)}\textsuperscript{*} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Language Tags}
		\label{tab:evaluation-disco-language-tags}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Aggregation}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-AGGREGATION-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-03 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-04 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-05 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-06 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-AGGREGATION-07 (!)}\textsuperscript{*} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Aggregation}
		\label{tab:evaluation-disco-aggregation}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
    \emph{DISCO-C-ALLOWED-VALUES-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-LITERAL-RANGES-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-INVERSE-FUNCTIONAL-PROPERTIES-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-INVERSE-FUNCTIONAL-PROPERTIES-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-CLASS-SPECIFIC-PROPERTY-RANGE-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-MEMBERSHIP-IN-CONTROLLED-VOCABULARIES-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & \ding{55} \\
		\emph{DISCO-C-LITERAL-VALUE-COMPARISON-01}\textsuperscript{***} & $\checkmark$ & 1,299 & $\checkmark$ \\
		\emph{DISCO-C-CONTEXT-SPECIFIC-VALID-PROPERTIES-01}\textsuperscript{*} & 21,038 & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-DATA-PROPERTY-FACETS-01}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-DATA-PROPERTY-FACETS-02}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (1)}
		\label{tab:evaluation-disco-disco-constraints-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-VALUE-IS-VALID-FOR-DATATYPE-01}\textsuperscript{***} & 30 & 6,932 & $\checkmark$ \\
		\emph{DISCO-C-VALUE-IS-VALID-FOR-DATATYPE-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DISCO-C-SUBSUMPTION-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-CLASS-EQUIVALENCE-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-SUB-PROPERTIES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-PROPERTY-DOMAIN-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-PROPERTY-RANGES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-INVERSE-OBJECT-PROPERTIES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-INVERSE-OBJECT-PROPERTIES-02 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-INVERSE-OBJECT-PROPERTIES-03 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DISJOINT-PROPERTIES-01 (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (2)}
		\label{tab:evaluation-disco-disco-constraints-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-ASYMMETRIC-OBJECT-PROPERTIES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-IRREFLEXIVE-OBJECT-PROPERTIES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-CLASS-SPECIFIC-IRREFLEXIVE-OBJECT-PROPERTIES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-CLASS-SPECIFIC-IRREFLEXIVE-OBJECT-PROPERTIES-02 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-DISJOINT-CLASSES-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-EQUIVALENT-PROPERTIES-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-LITERAL-PATTERN-MATCHING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-DISJUNCTION-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-UNIVERSAL-QUANTIFICATIONS-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01 (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (3)}
		\label{tab:evaluation-disco-disco-constraints-3}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-MAXIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-CONTEXT-SPECIFIC-EXCLUSIVE-OR-OF-PROPERTY-GROUPS-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-IRI-PATTERN-MATCHING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-ORDERING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-ORDERING-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-ORDERING-03 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-STRING-OPERATIONS-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-CONTEXT-SPECIFIC-VALID-CLASSES-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-CONTEXT-SPECIFIC-VALID-PROPERTIES-01 (!)}\textsuperscript{*} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (4)}
		\label{tab:evaluation-disco-disco-constraints-4}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-DEFAULT-VALUES-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-WHITESPACE-HANDLING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-HTML-HANDLING-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-HTML-HANDLING-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-RECOMMENDED-PROPERTIES-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-HANDLE-RDF-COLLECTIONS-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-HANDLE-RDF-COLLECTIONS-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-USE-SUB-SUPER-RELATIONS-IN-VALIDATION-01 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-USE-SUB-SUPER-RELATIONS-IN-VALIDATION-02 (!)}\textsuperscript{*} \\
		\emph{DISCO-C-STRUCTURE-01 (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (5)}
		\label{tab:evaluation-disco-disco-constraints-5}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccc@{}}
           & \multicolumn{3}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-4}
    \\       \textbf{Disco Constraints}
           & \rot{\emph{Missy}}
           & \rot{\emph{DwB}}
           & \rot{\emph{DDA-SND}}
    \\ \midrule
		\emph{DISCO-C-VOCABULARY-01 (!)}\textsuperscript{***} \\
		\emph{DISCO-C-HTTP-URI-SCHEME-VIOLATION (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Disco Data Sets - Disco Constraints (6)}
		\label{tab:evaluation-disco-disco-constraints-6}
    \end{center}
\end{table}

\section{Aggregated Metadata (Data Cube)}

\subsection{Evaluation Results}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Evaluation Criteria} & \textbf{Counts} \\		
      \hline
		\emph{Validated Triples} & 3,775,983,610 \\
		\emph{Validated Data Sets} & 9,990 \\
		\hline
		\emph{Constraint Violations} & 45,635,846 \\
		\emph{Constraint Violations ($\mathcal{SL}_{0}$)} & 0 $\equiv$ 0\% \\
		\emph{Constraint Violations ($\mathcal{SL}_{1}$)} & 45,520,613 $\equiv$ 99.75\% \\
    \emph{Constraint Violations ($\mathcal{SL}_{2}$)} & 115,233 $\equiv$ 0.25\% \\
		\hline
		\emph{Constraint (Most Constraint Violations)} & \emph{DATA-MODEL-CONSISTENCY-05} (45,514,102) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{0}$))} & - \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{1}$))} & \emph{DATA-MODEL-CONSISTENCY-05} (45,514,102) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{2}$))} & \emph{MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-02} \\
		                                                                    & (1,556) \\
		\hline
		\emph{Constraint Types} & 20 (7\textbar 13) \\
		\emph{Constraint Types} ($\mathcal{C}_{C}$) & 5 $\equiv$ 25 \% \\
		\emph{Constraint Types} ($\mathcal{C}_{T}$) & 15 $\equiv$ 75 \% \\
		\hline
		\emph{Constraint Types (Most Constraints)} & 1. Data Model Consistency: 11 $\equiv$ 31.4\% (10\textbar 1) \\
		                                           & 2. Existential Quantifications: 4 $\equiv$ 11.4\% (4\textbar 0) \\
		\emph{Constraint Type (Most Constraints ($\mathcal{SL}_{2}$))} & Data Model Consistency: 8 $\equiv$ 22.9\% (7\textbar 1) \\
		\hline
		\emph{Constraints} & 35 (20\textbar 15) \\
		\emph{Constraints} ($\mathcal{C}_{C}$) & 16 (12\textbar 4) \\
		\emph{Constraints} ($\mathcal{C}_{T}$) & 19 (8\textbar 11) \\
		\emph{Constraints ($\mathcal{SL}_{0}$)} & 4 $\equiv$ 11.4\% (0\textbar 4) \\
		\emph{Constraints ($\mathcal{SL}_{1}$)} & 3 $\equiv$ 8.6\% (3\textbar 0) \\
		\emph{Constraints ($\mathcal{SL}_{2}$)} & 28 $\equiv$ 80\% (17\textbar 11) \\
		\end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Evaluation Results}
		\label{tab:data-cube-evaluation-results}
\end{table}

\subsection{Data Sets Overview}

\begin{itemize}
  \item overview over Data Cube data sets\footnote{\url{http://270a.info/}; \url{http://datahub.io/de/dataset?tags=format-qb}} 
  \item \url{http://ontologycentral.com/}
\end{itemize}

%\begin{table}
	%\centering
		%\begin{tabular}{l|c|c|c|c|c|c}
      %\textbf{Data Set} & \textbf{triples} & \textbf{data sets} & \textbf{data structure definitions} & \textbf{observations} & \textbf{slices} & \textbf{dimensions} \\		
      %\hline
      %European Central Bank (ECB)\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} &  &  &  &  &  &  \\
			%UNESCO Institute for Statistics (UIS)\tablefootnote{\url{http://www.uis.unesco.org/Pages/default.aspx}} &  &  &  &  &  &  \\
			%International Monetary Fund (IMF)\tablefootnote{\url{http://www.imf.org/external/index.htm}} &  &  &  &  &  &  \\
			%Bundesamt für Statistik - Swiss Federal Statistics (BFS)\tablefootnote{\url{http://www.bfs.admin.ch/}} &  &  &  &  &  &  \\
			%Food and Agriculture Organization of the United Nations (FAO)\tablefootnote{\url{http://www.fao.org/home/en/}} &  &  &  &  &  &  \\
			%World Bank (WB)\tablefootnote{\url{http://www.worldbank.org/}} &  &  &  &  &  &  \\
			%Federal Reserve Board (FRB)\tablefootnote{\url{http://www.federalreserve.gov/}} &  &  &  &  &  &  \\
			%Transparency International (TI)\tablefootnote{\url{http://www.transparency.org/}} &  &  &  &  &  &  \\
			%Organisation for Economic Co-operation and Development (OECD)\tablefootnote{\url{http://www.oecd.org/}} &  &  &  &  &  &  \\
			%Bank for International Settlements (BIS)\tablefootnote{\url{http://www.bis.org/}} &  &  &  &  &  &  \\
			%Australian Bureau of Statistics (ABS)\tablefootnote{\url{http://abs.gov.au/}} &  &  &  &  &  &  \\
		%\end{tabular}
	%\caption{Data Cube Data Sets}
	%\label{tab:data-cube-data-sets}
%\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Data Cube Data Sets} \\		
      \hline
    \emph{ECB} & \emph{European Central Bank}\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} \\
		\emph{UIS} & \emph{UNESCO Institute for Statistics}\tablefootnote{\url{http://www.uis.unesco.org/Pages/default.aspx}} \\
		\emph{IMF} & \emph{International Monetary Fund}\tablefootnote{\url{http://www.imf.org/external/index.htm}} \\
		\emph{BFS} & \emph{Bundesamt für Statistik - Swiss Federal Statistics}\tablefootnote{\url{http://www.bfs.admin.ch/}} \\
		\emph{FAO} & \emph{Food and Agriculture Organization of the United Nations}\tablefootnote{\url{http://www.fao.org/home/en/}} \\
		\emph{WB} & \emph{World Bank}\tablefootnote{\url{http://www.worldbank.org/}} \\
		\emph{FRB} & \emph{Federal Reserve Board}\tablefootnote{\url{http://www.federalreserve.gov/}} \\
		\emph{TI} & \emph{Transparency International}\tablefootnote{\url{http://www.transparency.org/}} \\
		\emph{OECD} & \emph{Organisation for Economic Co-operation and Development}\tablefootnote{\url{http://www.oecd.org/}} \\
		\emph{BIS} & \emph{Bank for International Settlements}\tablefootnote{\url{http://www.bis.org/}} \\
		\emph{ABS} & \emph{Australian Bureau of Statistics}\tablefootnote{\url{http://abs.gov.au/}} \\
		\emph{IEEE-VIS} & \emph{IEEE VIS Source Data} \\
		\emph{ACORN-SAT} & \emph{Australian Climate Observations Reference Network - Surface Air Temperature Dataset} \\
		\emph{HDP} & \emph{HealthData.gov Platform (HDP) on the Semantic Web} \\
		\emph{Eurostat} & \emph{The Eurostat Linked Data} (SPARQL endpoint unavailable) \\
		\emph{Asturias} & \emph{Nomenclator Asturias} (SPARQL endpoint unavailable!) \\
		\emph{ISTAT} & \emph{ISTAT Immigration (LinkedOpenData.it)} (SPARQL endpoint unavailable) \\
		\emph{ICANE} & \emph{Statistical Office of Cantabria (Instituto Cántabro de Estadística, ICANE)} \\ 
		      & (SPARQL endpoint unavailable) \\
		\emph{EE-2009} & \emph{European Election Results 2009} (SPARQL endpoint unavailable) \\
		\emph{EU-B} & \emph{Standard Eurobarometer} (SPARQL endpoint unavailable) \\
		\emph{ECB-S} & \emph{European Central Bank Statistics (PublicData.eu)} (SPARQL endpoint unavailable) \\
		\emph{CPV-2008} & \emph{Common Procurement Vocabulary (CPV) 2008} (SPARQL endpoint unavailable) \\
		\emph{CPV-2003} & \emph{Common Procurement Vocabulary (CPV) 2003} (SPARQL endpoint unavailable) \\
		%GHI-2011 & Global Hunger Index (GHI) 2011 (dump) \\
		\end{tabular}
	\caption{Data Cube Data Sets Abbreviations}
	\label{tab:data-cube-data-sets-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{5}{c}{\textbf{Counts}}
    \\  \cmidrule{2-6}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{qb:DataSet}}
           & \textbf{\rot{qb:DataStructureDefinition}}
           & \textbf{\rot{qb:Observation}}
           & \textbf{\rot{qb:Slice}}
    \\ \midrule
    \emph{ECB} & 468,899,474 & 55 & 46 & $>$11,000,000 & 428,698 \\
		\emph{UIS} & 10,400,534 & 5 & 5 & 1,437,651 & 0 &  \\
		\emph{IMF} & 35,688,446 & 4 & 8 & 3,603,719 & 0 &  \\
		\emph{BFS} & 1,533,743 & 0 & 0 & 8 & 0 &  \\
		\emph{FAO} & 53,000,000 & 10 & 10 & $>$7,100,000 & 0 \\
		\emph{WB} & 174,006,552 & 9,466 & 59 & $>$17,000,000 & 0 \\
		\emph{FRB} & 185,266,900 & 49 & 98 & $>$9,500,000 & 0 \\
		\emph{TI} & 52,233 & 6 & 6 & 3,928 & 0 \\
		\emph{OECD} & 304,995,160 & 136 & 140 & $>$12,000,000 & 0 \\
		\emph{BIS} & 54,197,482 & 6 & 12 & 3,606,466 & 47,914 \\
		\emph{ABS} & 2,357,400,000 & 253 & 257 & $>$11,000,000 & 0  \\
		\emph{IEEE-VIS} & 19,935,340 & 0 & 0 & 1,350 & 0 \\
		\emph{ACORN-SAT} & 98,381,319 & 0 & 4 & 0 & 0 \\
		\emph{HDP} & 12,226,427 & 0 & 0 & 0 & 0 \\
		%Eurostat & 8,000,000,000 &  &  &  &  &  \\
		%Asturias & 4,508,050 &  &  &  &  &  \\
		%ISTAT & 3,024,396 &  &  &  &  &  \\
		%ICANE & 52,412 &  &  &  &  &  \\
		%EE-2009 & 3,165 &  &  &  &  &  \\
		%EU-B & 1,193,494 &  &  &  &  &  \\
		%ECB-S & 10,000,000 &  &  &  &  &  \\
		%CPV-2008 & 803,311 &  &  &  &  &  \\
		%CPV-2003 & 546,135 &  &  &  &  &  \\
		%GHI-2011 & 19,369 (dump) &  &  &  &  &  \\
		\hline
		\textbf{Total} & 3,775,983,610 & 9,990 & & & & \\
    \bottomrule
    \end{tabular}
    \caption{Data Cube Data Sets Overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Data Sets} & \textbf{SPARQL Endpoints} \\		
      \hline
      \emph{ECB} & \url{http://ecb.270a.info/sparql} \\
			\emph{UIS} & \url{http://uis.270a.info/sparql} \\
			\emph{IMF} & \url{http://imf.270a.info/sparql} \\
			\emph{BFS} & \url{http://bfs.270a.info/sparql} \\
			\emph{FAO} & \url{http://fao.270a.info/sparql} \\
			\emph{WB} & \url{http://worldbank.270a.info/sparql} \\
			\emph{FRB} & \url{http://frb.270a.info/sparql} \\
			\emph{TI} & \url{http://transparency.270a.info/sparql} \\
			\emph{OECD} & \url{http://oecd.270a.info/sparql} \\
			\emph{BIS} & \url{http://bis.270a.info/sparql} \\
			\emph{ABS} & \url{http://abs.270a.info/sparql} \\
			\emph{ACORN-SAT} & \url{http://lab.environment.data.gov.au/sparql} \\
			\emph{HDP} & \url{http://healthdata.tw.rpi.edu/sparql} \\
		\end{tabular}
	\caption{Data Cube SPARQL Endpoints}
	\label{tab:data-cube-sparql-endpoints}
\end{table}

\subsection{Detailed Evaluation}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccc@{}}
           & \multicolumn{7}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-8}
    \\       \textbf{Data Model Consistency}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
    \\ \midrule
    \emph{DATA-MODEL-CONSISTENCY-01}\textsuperscript{**} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-02}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-03}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-04}\textsuperscript{***} & $\checkmark$ (6) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 14,372 \\
		\emph{DATA-MODEL-CONSISTENCY-05}\textsuperscript{**} & 1,198,352 (50) & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ & 16,175,814 (42) \\
		\emph{DATA-MODEL-CONSISTENCY-06}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-07}\textsuperscript{***} & $\checkmark$ (9) & $\checkmark$ & 99,091 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ (1) \\
		\emph{DATA-MODEL-CONSISTENCY-08}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-09}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-10}\textsuperscript{***} (!) & - & - & - & - & - & - & - \\
		\emph{DATA-MODEL-CONSISTENCY-11}\textsuperscript{**} & 6,511 (10) & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Data Model Consistency (1)}
		\label{evaluation-of-data-cube-data-sets-data-model-consistency-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccc@{}}
           & \multicolumn{7}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-8}
    \\       \textbf{Data Model Consistency}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
    \emph{DATA-MODEL-CONSISTENCY-01}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 8 & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-03}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \emph{DATA-MODEL-CONSISTENCY-04}\textsuperscript{***} & $\checkmark$ & $\checkmark$& $\checkmark$ & $\checkmark$ (6) & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-05}\textsuperscript{**} & $\checkmark$ & 21,142,838 (116) & \ding{55} &  6,997,098 (246) & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-06}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-07}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ (8) & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-08}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-09}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{DATA-MODEL-CONSISTENCY-10}\textsuperscript{***} (!) & - & - & - & - & - & - & - \\
		\emph{DATA-MODEL-CONSISTENCY-11}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Data Model Consistency (2)}
		\label{evaluation-of-data-cube-data-sets-data-model-consistency-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccccc@{}}
           & \multicolumn{14}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-15}
    \\       \textbf{Existential Quantifications}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
		\emph{EXISTENTIAL-QUANTIFICATIONS-01}\textsuperscript{***} & 9 & $\checkmark$ & 11 & 7 & 8 & 77 & 8 & 9 & 7 & 8 & 7 & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXISTENTIAL-QUANTIFICATIONS-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXISTENTIAL-QUANTIFICATIONS-03}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 59 & $\checkmark$ & 6 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 4 & $\checkmark$ \\
		\emph{EXISTENTIAL-QUANTIFICATIONS-04}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Existential Quantifications}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccc@{}}
           & \multicolumn{10}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-11}
    \\       \textbf{Cardinality Restrictions}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
    \\ \midrule
		\emph{MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01 (!)}\textsuperscript{***} & - & - & - & - & - & - & - & - & - & - \\
		\emph{MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-02}\textsuperscript{***} & \ding{55} & 118 & 8 & 8 & 30 & $\checkmark$ & 30 & $\checkmark$ & \ding{55} & 12 \\
		\emph{MAXIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXACT-UNQUALIFIED-CARDINALITY-RESTRICTIONS-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Cardinality Restrictions (1)}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Cardinality Restrictions}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
		\emph{MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01 (!)}\textsuperscript{***} & - & - & - & - \\
		\emph{MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-02}\textsuperscript{***} & \ding{55} & 1,350 & $\checkmark$ & $\checkmark$ \\
		\emph{MAXIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-01}\textsuperscript{***} & $\checkmark$ (2) & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXACT-UNQUALIFIED-CARDINALITY-RESTRICTIONS-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Cardinality Restrictions (2)}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccccc@{}}
           & \multicolumn{14}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-15}
    \\       \textbf{Structure}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
		\emph{STRUCTURE-01}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-02}\textsuperscript{***} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Structure}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccccc@{}}
           & \multicolumn{14}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-15}
    \\       \textbf{Constraints}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
		\emph{PROPERTY-DOMAIN-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{PROPERTY-RANGES-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{DISJOINT-PROPERTIES-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{DISJOINT-CLASSES-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{EQUIVALENT-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{UNIVERSAL-QUANTIFICATIONS-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{MEMBERSHIP-IN-CONTROLLED-VOCABULARIES-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{CONTEXT-SPECIFIC-VALID-CLASSES-01 (!)}\textsuperscript{*} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{CONTEXT-SPECIFIC-VALID-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{RECOMMENDED-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{VALUE-IS-VALID-FOR-DATATYPE-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
		\emph{VOCABULARY-01 (!)}\textsuperscript{***} &  &  &  &  &  &  &  &  &  &  &  &  \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Constraints (1)}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccccc@{}}
           & \multicolumn{14}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-15}
    \\       \textbf{Constraints}
           & \rot{\emph{ECB}}
           & \rot{\emph{UIS}}
           & \rot{\emph{IMF}}
           & \rot{\emph{BFS}}
           & \rot{\emph{FAO}}
					 & \rot{\emph{WB}}
					 & \rot{\emph{FRB}}
					 & \rot{\emph{TI}}
					 & \rot{\emph{OECD}}
					 & \rot{\emph{BIS}}
					 & \rot{\emph{ABS}}
					 & \rot{\emph{IEEE-VIS}}
					 & \rot{\emph{ACORN-SAT}}
					 & \rot{\emph{HDP}}
    \\ \midrule
		\emph{HTTP-URI-SCHEME-VIOLATION (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Evaluation of Data Cube Data Sets - Constraints (2)}
    \end{center}
\end{table}

\section{Thesauri (SKOS)}

\subsection{Evaluation Results}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Evaluation Criteria} & \textbf{Counts} \\		
      \hline
		\emph{Validated Triples} & \textcolor{red}{XXXXX} \\
		\emph{Validated Data Sets} & \textcolor{red}{XXXXX} \\
		\hline
		\emph{Constraint Violations} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{0}$)} & \textcolor{red}{XXXXX} \\
		\emph{Constraint Violations ($\mathcal{SL}_{1}$)} & \textcolor{red}{XXXXX} $\equiv$ \textcolor{red}{XXXXX}\% \\
    \emph{Constraint Violations ($\mathcal{SL}_{2}$)} & \textcolor{red}{XXXXX} \\
		\hline
		\emph{Constraint (Most Constraint Violations)} & \emph{LANGUAGE-TAG-CARDINALITY-01} (2,508,903) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{0}$))} & \emph{LABELING-AND-DOCUMENTATION-06} (1,022,362) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{1}$))} & \emph{LANGUAGE-TAG-CARDINALITY-01} (2,508,903) \\
		\emph{Constraint (Most Constraint Violations ($\mathcal{SL}_{2}$))} & - \\
		\hline
		\emph{Constraint Types} & 14 (4\textbar 10) \\
		\emph{Constraint Types} ($\mathcal{C}_{C}$) & 5 $\equiv$ 35.7\% \\
		\emph{Constraint Types} ($\mathcal{C}_{T}$) & 9 $\equiv$ 64.3\% \\
		\hline
		\emph{Constraint Types (Most Constraints)} & 1. Structure: 10 $\equiv$ XX\% (8\textbar 2) \\
		                                           & 2. Labeling and Documentation: 6 $\equiv$ XX\% (5\textbar 1) \\
																							 & 3. Language Tag Cardinality: 4 $\equiv$ XX\% (4\textbar 0) \\
		\emph{Constraint Type (Most Constraints ($\mathcal{SL}_{2}$))} & Structure: 1 (0\textbar 1) \\
		\hline
		\emph{Constraints} & 35 (17\textbar 18) \\
		\emph{Constraints} ($\mathcal{C}_{C}$) & 21 (13\textbar 8) \\
		\emph{Constraints} ($\mathcal{C}_{T}$) & 14 (4\textbar 10) \\
		\emph{Constraints ($\mathcal{SL}_{0}$)} & 21 $\equiv$ 60\% (12\textbar 9) \\
		\emph{Constraints ($\mathcal{SL}_{1}$)} & 5 $\equiv$ 14.3\% (5\textbar 0) \\
		\emph{Constraints ($\mathcal{SL}_{2}$)} & 9 $\equiv$ 25.7\% (0\textbar 9) \\
		\end{tabular}
    \caption{Evaluation of Thesauri Data Sets - Evaluation Results}
		\label{tab:thesauri-evaluation-results}
\end{table}

\subsection{Data Sets Overview}

\begin{itemize}
  \item overview over SKOS data sets\footnote{\url{http://datahub.io/de/dataset?tags=format-skos}}
	\item overview over thesauri\footnote{\url{http://datahub.io/de/dataset?tags=thesaurus}} (table \ref{tab:thesauri-overview})
\end{itemize}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Thesauri} \\		
      \hline
    \emph{TheSoz} & \emph{Thesaurus for the Social Sciences}\tablefootnote{\url{http://www.ecb.europa.eu/home/html/index.en.html}} \\
	  \emph{STW} & \emph{Thesaurus for Economics}\tablefootnote{\url{http://zbw.eu/stw/versions/latest/about}} \\
	  \emph{AGROVOC} & \emph{AGROVOC Multilingual agricultural thesaurus}\tablefootnote{\url{http://202.45.139.84:10035/catalogs/fao/repositories/agrovoc}} \\
		\emph{UNESCO} & \emph{UNESCO Thesaurus}\tablefootnote{\url{http://skos.um.es/sparql/}} \\
		\emph{TGN} & \emph{The Getty Thesaurus of Geographic Names}\tablefootnote{\url{http://vocab.getty.edu/sparql}} \\
		\emph{EARTh} & \emph{Environmental Applications Reference Thesaurus}\tablefootnote{\url{http://linkeddata.ge.imati.cnr.it/resource/EARTh/}} \\
		\emph{ODT} & \emph{Open Data Thesaurus}\tablefootnote{\url{http://vocabulary.semantic-web.at/PoolParty/wiki/OpenData}} \\
		\emph{SLD} & \emph{Spanish Linguistic Datasets}\tablefootnote{\url{http://linguistic.linkeddata.es}} \\
		\emph{SSWT} & \emph{Social Semantic Web Thesaurus}\tablefootnote{\url{http://vocabulary.semantic-web.at/PoolParty/wiki/semweb}} \\
		\emph{GBA-GU} & \emph{Thesaurus of the Geological Survey of Austria (GBA) - Geology Unit}\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		\emph{GBA-GTS} & \emph{Thesaurus of the Geological Survey of Austria (GBA) - Geologic Time Scale}\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		\emph{GBA-L} & \emph{Thesaurus of the Geological Survey of Austria (GBA) - Lithology}\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		\emph{GBA-LU} & \emph{Thesaurus of the Geological Survey of Austria (GBA) - Lithotectonic Unit}\tablefootnote{\url{http://resource.geolba.ac.at/}} \\
		\emph{GEMET} & \emph{GEneral Multilingual Environmental Thesaurus}\tablefootnote{\url{http://www.eionet.europa.eu/gemet/}} \\
		\emph{EuroVoc} & \emph{EuroVoc}\tablefootnote{\url{http://open-data.europa.eu/de/data/dataset/eurovoc}} \\
		\emph{CECCT} & \emph{Clean Energy and Climate Change Thesaurus}\tablefootnote{\url{http://data.reegle.info/thesaurus/guide}} \\
		\end{tabular}
	\caption{Thesauri Abbreviations}
	\label{tab:thesauri-abbreviations}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{7}{c}{\textbf{Counts}}
    \\  \cmidrule{2-8}
    \\       \textbf{Thesauri}
           & \textbf{\rot{triples}}
           & \textbf{\rot{skos:ConceptScheme}}
           & \textbf{\rot{sko:Concept}}
           & \textbf{\rot{skos:broader}}
           & \textbf{\rot{skos:narrower}}
					 & \textbf{\rot{skos:hasTopConcept}}
					 & \textbf{\rot{skos:inScheme}}
    \\ \midrule
    \emph{TheSoz} & 439,153 & 1 & 8,426 & 13,705 & 13,706 & 0 & 48,529 \\
	  \emph{STW} & 221,668 & 1 & 13,468 & 13,732 & 13732 & 7 & 13,180 \\
	  \emph{AGROVOC} & 6,080,477 & 1 & 32,310 & 33,507 & 33,507 & 25 & 32,310 \\
		\emph{UNESCO} & 288,346 & 9 & 26,714 & 20,028 & 20,028 & 607 & 32,009 \\
		\emph{TGN} & 16,112,321 & 8 & 2,898,775 & 0 & 0 & 0 & 1,453,767\\
		\emph{EARTh} & 9,287,364 & 11 & 295,375 & 288,208 & 93,827 & 479 & 295,376 \\
		\emph{ODT} & 3,290 & 6 & 108 & 93 & 93 & 30 & 0 \\
		\emph{SLD} & 7,629,211 & 0 & 31,195 & 0 & 0 & 0 & 0 \\
		\emph{SSWT} & 64,698 & 9 & 2,127 & 2,300 & 2,301 & 38 & 0 \\
		\emph{GBA-GU} & 25,718 & 3 & 878 & 1,005 & 1,005 & 14 & 0 \\
		\emph{GBA-GTS} & 7,875 & 3 & 213 & 208 & 208 & 5 & 0 \\
		\emph{GBA-L} & 9,317 & 1 & 249 & 249 & 249 & 4 & 0 \\
		\emph{GBA-LU} & 9,504 & 3 & 364 & 359 & 359 & 7 & 0 \\
		\emph{GEMET} & 372,889,229 & 3,680 & 414,659 & 62,193 & 21,685 & 30,806 & 409,290 \\
		\emph{EuroVoc} & 64,477,774 & 439 & 79,557 & 6,922 & 0 & 532 & 14,428 \\
		\emph{CECCT} & 191,336 & 3 & 3,419 & 3,761 & 3,762 & 28 & 0 \\
		\hline
		\textbf{Total} & 477,737,281 & 4,178 & & & & \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Overview}
		\label{tab:thesauri-overview}
    \end{center}
\end{table}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Thesauri} & \textbf{SPARQL Endpoints} \\		
      \hline
      \emph{TheSoz} & \url{http://lod.gesis.org/thesoz/sparql} \\
			\emph{STW} & \url{http://zbw.eu/beta/sparql/stw/query} \\
			\emph{AGROVOC} & \url{http://202.45.139.84:10035/catalogs/fao/repositories/agrovoc} \\
			\emph{UNESCO} & \url{http://skos.um.es/sparql/} \\
			\emph{TGN} & \url{http://vocab.getty.edu/} \\
			\emph{EARTh} & \url{http://linkeddata.ge.imati.cnr.it:8890/sparql} \\
			\emph{ODT} & \url{http://vocabulary.semantic-web.at/PoolParty/sparql/OpenData} \\
			\emph{SLD} & \url{http://linguistic.linkeddata.es/sparql} \\
			\emph{SSWT} & \url{http://vocabulary.semantic-web.at/PoolParty/sparql/semweb} \\
			\emph{GBA-GU} & \url{http://resource.geolba.ac.at/PoolParty/sparql/GeologicUnit} \\
			\emph{GBA-GTS} & \url{http://resource.geolba.ac.at/PoolParty/sparql/GeologicTimeScale} \\
			\emph{GBA-L} & \url{http://resource.geolba.ac.at/PoolParty/sparql/lithology} \\
			\emph{GBA-LU} & \url{http://resource.geolba.ac.at/PoolParty/sparql/tectonicunit} \\
			\emph{GEMET} & \url{http://semantic.eea.europa.eu/sparql} \\
			\emph{EuroVoc} & \url{http://open-data.europa.eu/de/linked-data} \\
			\emph{CECCT} & \url{http://poolparty.reegle.info/PoolParty/sparql/glossary}
		\end{tabular}
	\caption{Thesauri SPARQL Endpoints}
	\label{tab:thesauri-sparql-endpoints}
\end{table}

\subsection{Detailed Evaluation}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Data Model Consistency}
           & \emph{\rot{TheSoz}}
           & \emph{\rot{STW}}
           & \emph{\rot{AGROVOC}}
					 & \emph{\rot{TGN}}
           & \emph{\rot{UNESCO}}
					 & \emph{\rot{ODT}}
					 & \emph{\rot{SSWT}}
					 & \emph{\rot{GBA-GU}}
					 & \emph{\rot{GBA-GTS}}
					 & \emph{\rot{GBA-L}}
					 & \emph{\rot{GBA-LU}}
					 & \emph{\rot{CECCT}}
    \\ \midrule
    \emph{DATA-MODEL-CONSISTENCY-01 (!)}\textsuperscript{*} &  &  &  &  &  & \\
    \emph{DATA-MODEL-CONSISTENCY-02 (!)}\textsuperscript{*} &  &  &  &  &  & \\
    \emph{DATA-MODEL-CONSISTENCY-03 (!)}\textsuperscript{*} &  &  &  &  &  & \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Data Model Consistency (1)}
		\label{tab:thesauri-evaluation-data-model-consistency-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Data Model Consistency}
					 & \emph{\rot{EARTh}}
					 & \emph{\rot{GEMET}}
					 & \emph{\rot{EuroVoc}}
					 & \emph{\rot{SLD}}
    \\ \midrule
    \emph{DATA-MODEL-CONSISTENCY-01 (!)}\textsuperscript{*} \\
    \emph{DATA-MODEL-CONSISTENCY-02 (!)}\textsuperscript{*} \\
    \emph{DATA-MODEL-CONSISTENCY-03 (!)}\textsuperscript{*} \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Data Model Consistency (2)}
		\label{tab:thesauri-evaluation-data-model-consistency-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Labeling and Documentation}
           & \emph{\rot{TheSoz}}
           & \emph{\rot{STW}}
           & \emph{\rot{AGROVOC}}
					 & \emph{\rot{TGN}}
           & \emph{\rot{UNESCO}}
					 & \emph{\rot{ODT}}
					 & \emph{\rot{SSWT}}
					 & \emph{\rot{GBA-GU}}
					 & \emph{\rot{GBA-GTS}}
					 & \emph{\rot{GBA-L}}
					 & \emph{\rot{GBA-LU}}
					 & \emph{\rot{CECCT}}
    \\ \midrule
		\emph{LABELING-AND-DOCUMENTATION-01}\textsuperscript{*} & 8,426 & 11,508 & 19,829 & 1,110 & \ding{55} & 36 & 1,475 & 5 & 2 & $\checkmark$ & 107 & 486 \\
		\emph{LABELING-AND-DOCUMENTATION-02}\textsuperscript{*} & $>$1 & \ding{55} & $>$100 & 287 & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{LABELING-AND-DOCUMENTATION-03}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & 1 & 14,114 & \ding{55} & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ & $\checkmark$ & 1 & $\checkmark$ \\
		\emph{LABELING-AND-DOCUMENTATION-04 (!)}\textsuperscript{*} &  &  &  &  &  & & &  \\
		\emph{LABELING-AND-DOCUMENTATION-05}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & 4 & $\checkmark$ & 1 & 2 & 2 & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & 7 \\
		\emph{LABELING-AND-DOCUMENTATION-06}\textsuperscript{*} & 975,340 & $\checkmark$ & $\checkmark$ & 2 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Labeling and Documentation (1)}
		\label{tab:thesauri-evaluation-labeling-and-documentation-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Labeling and Documentation}
					 & \emph{\rot{EARTh}}
					 & \emph{\rot{GEMET}}
					 & \emph{\rot{EuroVoc}}
					 & \emph{\rot{SLD}}
    \\ \midrule
		\emph{LABELING-AND-DOCUMENTATION-01}\textsuperscript{*}& 264,687 & \ding{55} & 54,911 & 31,195 \\
		\emph{LABELING-AND-DOCUMENTATION-02}\textsuperscript{*} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
		\emph{LABELING-AND-DOCUMENTATION-03}\textsuperscript{*} & 2 & \ding{55} & 55,556 & 31,195 \\
		\emph{LABELING-AND-DOCUMENTATION-04 (!)}\textsuperscript{*} & \\
		\emph{LABELING-AND-DOCUMENTATION-05}\textsuperscript{*} & 39 & \ding{55} & \ding{55} & 978 \\
		\emph{LABELING-AND-DOCUMENTATION-06}\textsuperscript{*} &  302 & 46,718 & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Labeling and Documentation (2)}
		\label{tab:thesauri-evaluation-labeling-and-documentation-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Structure}
           & \emph{\rot{TheSoz}}
           & \emph{\rot{STW}}
           & \emph{\rot{AGROVOC}}
					 & \emph{\rot{TGN}}
           & \emph{\rot{UNESCO}}
					 & \emph{\rot{ODT}}
					 & \emph{\rot{SSWT}}
					 & \emph{\rot{GBA-GU}}
					 & \emph{\rot{GBA-GTS}}
					 & \emph{\rot{GBA-L}}
					 & \emph{\rot{GBA-LU}}
					 & \emph{\rot{CECCT}}
    \\ \midrule
		\emph{STRUCTURE-01}\textsuperscript{**} & 1 & 1,074 & $\checkmark$ & $\checkmark$ & 1 & 5 & 1 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-02 (!)}\textsuperscript{*} &  &  &  &  &  & \\
		\emph{STRUCTURE-03}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 84 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-04}\textsuperscript{*} & 2,906 & 8,046 & 726 & $\checkmark$ & 3,840 & 12 & 124 & 84 & 256 & 68 & 22 & 2,422 \\
		\emph{STRUCTURE-05}\textsuperscript{*} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & \ding{55} & 90 & 5,150 & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & 9,864 \\
		\emph{STRUCTURE-06}\textsuperscript{*} & 1,457 & 37 & $\checkmark$ & $\checkmark$ & \ding{55} & $\checkmark$ & 4 & 1 & 1 & 64 & $\checkmark$ & 136 \\
		\emph{STRUCTURE-07}\textsuperscript{**} & 40 & 5,370 & $\checkmark$ & $\checkmark$ & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-08 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{STRUCTURE-09}\textsuperscript{*} & 7,897 & 19,844 & 99 & $\checkmark$ & 552 & 2 & 16 & 26 & $\checkmark$ & $\checkmark$ & $\checkmark$ & 82 \\
		\emph{STRUCTURE-10}\textsuperscript{**} & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Structure (1)}
		\label{tab:thesauri-evaluation-structure-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Structure}
					 & \emph{\rot{EARTh}}
					 & \emph{\rot{GEMET}}
					 & \emph{\rot{EuroVoc}}
					 & \emph{\rot{SLD}}
    \\ \midrule
		\emph{STRUCTURE-01}\textsuperscript{**} & 18,240 & \ding{55} & 55,757 & 31,195 \\
		\emph{STRUCTURE-02 (!)}\textsuperscript{*} & \\
		\emph{STRUCTURE-03}\textsuperscript{**} & 39 & 4,244 & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-04}\textsuperscript{*} & 11,286 & 74 & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-05}\textsuperscript{*} & $\checkmark$ & \ding{55} & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-06}\textsuperscript{*} & 239,346 & \ding{55} & 13,876 & $\checkmark$ \\
		\emph{STRUCTURE-07}\textsuperscript{**} & 110,015 & \ding{55} & 366,155 & 155,975 \\
		\emph{STRUCTURE-08 (!)}\textsuperscript{***} & \\
		\emph{STRUCTURE-09}\textsuperscript{*} & 107,195 & 32 & $\checkmark$ & $\checkmark$ \\
		\emph{STRUCTURE-10}\textsuperscript{**} & 27 & 2,122 & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Structure (2)}
		\label{tab:thesauri-evaluation-structure-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Language Tag Cardinality}
           & \emph{\rot{TheSoz}}
           & \emph{\rot{STW}}
           & \emph{\rot{AGROVOC}}
					 & \emph{\rot{TGN}}
           & \emph{\rot{UNESCO}}
					 & \emph{\rot{ODT}}
					 & \emph{\rot{SSWT}}
					 & \emph{\rot{GBA-GU}}
					 & \emph{\rot{GBA-GTS}}
					 & \emph{\rot{GBA-L}}
					 & \emph{\rot{GBA-LU}}
					 & \emph{\rot{CECCT}}
    \\ \midrule
		\emph{LANGUAGE-TAG-CARDINALITY-01}\textsuperscript{**} & 9,435 & 13,468 & 98,894 & $\checkmark$ &  & 541 & 10,147 & 5,117 & 2,061 & 1,742 & 2,272 & 15,550 \\
		\emph{LANGUAGE-TAG-CARDINALITY-02}\textsuperscript{*} & 8,222 & 36,936 & \ding{55} & $\checkmark$ &  & 265 & 3,627 & 2,212 & 635 & 631 & 1,253 & 9,607 \\
		\emph{LANGUAGE-TAG-CARDINALITY-03}\textsuperscript{*} & 8,222 & $\checkmark$ & 135 & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
		\emph{LANGUAGE-TAG-CARDINALITY-04}\textsuperscript{*} & $\checkmark$ & 476 & \ding{55} & 50 &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Language Tag Cardinality (1)}
		\label{tab:thesauri-evaluation-language-tag-cardinality-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Language Tag Cardinality}
					 & \emph{\rot{EARTh}}
					 & \emph{\rot{GEMET}}
					 & \emph{\rot{EuroVoc}}
					 & \emph{\rot{SLD}}
    \\ \midrule
		\emph{LANGUAGE-TAG-CARDINALITY-01}\textsuperscript{**} & \ding{55} & 2,318,895 & \ding{55} & 30,781 \\
		\emph{LANGUAGE-TAG-CARDINALITY-02}\textsuperscript{*} & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
		\emph{LANGUAGE-TAG-CARDINALITY-03}\textsuperscript{*} & 224,206 & \ding{55} & \ding{55} & 31,195 \\
		\emph{LANGUAGE-TAG-CARDINALITY-04}\textsuperscript{*} & \ding{55} & \ding{55} & $\checkmark$ & $\checkmark$ \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Language Tag Cardinality (2)}
		\label{tab:thesauri-evaluation-language-tag-cardinality-2}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccccccccccc@{}}
           & \multicolumn{12}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-13}
    \\       \textbf{Constraints}
           & \emph{\rot{TheSoz}}
           & \emph{\rot{STW}}
           & \emph{\rot{AGROVOC}}
					 & \emph{\rot{TGN}}
           & \emph{\rot{UNESCO}}
					 & \emph{\rot{ODT}}
					 & \emph{\rot{SSWT}}
					 & \emph{\rot{GBA-GU}}
					 & \emph{\rot{GBA-GTS}}
					 & \emph{\rot{GBA-L}}
					 & \emph{\rot{GBA-LU}}
					 & \emph{\rot{CECCT}}
    \\ \midrule
		\emph{PROPERTY-DOMAIN-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{PROPERTY-RANGES-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{DISJOINT-PROPERTIES-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{DISJOINT-PROPERTIES-02 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{DISJOINT-CLASSES-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{EQUIVALENT-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  & \\
		\emph{UNIVERSAL-QUANTIFICATIONS-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{CONTEXT-SPECIFIC-VALID-CLASSES-01 (!)}\textsuperscript{*} \\ &  &  &  &  &  & \\
		\emph{CONTEXT-SPECIFIC-VALID-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  & \\
		\emph{RECOMMENDED-PROPERTIES-01 (!)}\textsuperscript{*} &  &  &  &  &  & \\
		\emph{VOCABULARY-01 (!)}\textsuperscript{***} &  &  &  &  &  & \\
		\emph{HTTP-URI-SCHEME-VIOLATION (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Constraints (1)}
		\label{tab:thesauri-evaluation-constraints-1}
    \end{center}
\end{table}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lcccc@{}}
           & \multicolumn{4}{c}{\textbf{Data Sets}}
    \\  \cmidrule{2-5}
    \\       \textbf{Constraints}
					 & \emph{\rot{EARTh}}
					 & \emph{\rot{GEMET}}
					 & \emph{\rot{EuroVoc}}
					 & \emph{\rot{SLD}}
    \\ \midrule
		\emph{PROPERTY-DOMAIN-01 (!)}\textsuperscript{***}\\
		\emph{PROPERTY-RANGES-01 (!)}\textsuperscript{***}\\
		\emph{DISJOINT-PROPERTIES-01 (!)}\textsuperscript{***}\\
		\emph{DISJOINT-PROPERTIES-02 (!)}\textsuperscript{***}\\
		\emph{DISJOINT-CLASSES-01 (!)}\textsuperscript{***}\\
		\emph{EQUIVALENT-PROPERTIES-01 (!)}\textsuperscript{*}\\
		\emph{UNIVERSAL-QUANTIFICATIONS-01 (!)}\textsuperscript{***}\\
		\emph{CONTEXT-SPECIFIC-VALID-CLASSES-01 (!)}\textsuperscript{*} \\
		\emph{CONTEXT-SPECIFIC-VALID-PROPERTIES-01 (!)}\textsuperscript{*} \\
		\emph{RECOMMENDED-PROPERTIES-01 (!)}\textsuperscript{*} \\
		\emph{VOCABULARY-01 (!)}\textsuperscript{***} \\
		\emph{HTTP-URI-SCHEME-VIOLATION (!)}\textsuperscript{***} \\
    \bottomrule
    \end{tabular}
    \caption{Thesauri Evaluation - Constraints (2)}
		\label{tab:thesauri-evaluation-constraints-2}
    \end{center}
\end{table}

\section{Rectangular Data (PHDD)}

\tb{ToDo}

\subsection{Evaluation Results}

\subsection{Data Sets Overview}

\subsection{Detailed Evaluation}

\section{Statistical Classifications (XKOS)}

\tb{ToDo}

\subsection{Evaluation Results}

\subsection{Data Sets Overview}

\begin{table}[H]
	\centering
		\begin{tabular}{l|l}
      \textbf{Abbr.} & \textbf{Statistical Classifications} \\		
      \hline
    \emph{NAF} & \emph{Nomenclature d'activités française}\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} \\
		\emph{PCS} & \emph{Nomenclature des Professions et Catégories Socioprofessionnelles}\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} \\
		\emph{CJ} & \emph{Nomenclature des catégories juridiques}\tablefootnote{\url{http://rdf.insee.fr/codes/index.html}} \\
		\emph{ISIC} & \emph{}\tablefootnote{\url{}} \\
		\emph{ISCO} & \emph{}\tablefootnote{\url{}} \\
		\end{tabular}
	\caption{Statistical Classifications Abbreviations}
	\label{tab:statistical-classifications-abbreviations}
\end{table}

\begin{itemize}
	\item \emph{Nomenclature d'activités française (NAF)}: French classification expressed in XKOS. the French refinement of the NACE, because it has explanatory notes.
	\item \emph{Nomenclature des Professions et Catégories Socioprofessionnelles (PCS)}: French classification expressed in XKOS.
	\item \emph{Nomenclature des catégories juridiques (CJ)}: French classification expressed in XKOS.
	\item \emph{ISIC}: has explanatory notes too. 
	\item \emph{ISCO}
\end{itemize}

\begin{table}[H]
    \begin{center}
    \begin{tabular}{@{}lccccccccccc@{}}
           & \multicolumn{6}{c}{\textbf{Counts}}
    \\  \cmidrule{2-7}
    \\       \textbf{Data Sets}
           & \textbf{\rot{triples}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
           & \textbf{\rot{}}
					 & \textbf{\rot{}}
					 & \textbf{\rot{}}
    \\ \midrule
    \emph{NAF} &  &  &  &  &  & \\
		\emph{PCS} &  &  &  &  &  & \\
		\emph{CJ} &  &  &  &  &  & \\
		\emph{ISIC} &  &  &  &  &  & \\
		\emph{ISCO} &  &  &  &  &  & \\
		\hline
		\textbf{Total} & & & & & & \\
    \bottomrule
    \end{tabular}
    \caption{Statistical Classifications Overview}
		\label{tab:statistical-classifications-overview}
    \end{center}
\end{table}

\subsection{Detailed Evaluation}

\bibliography{../../literature/literature}{}
\bibliographystyle{plain}
\setcounter{tocdepth}{1}
%\listoftodos
\end{document}